---
title: "Cropland abandonment benefits more birds and mammals than it harms, but rarely compensates for habitat loss"
author: Christopher L. Crawford,$^a$$^*$, R. Alex Wiebe,$^{d}$ He Yin,$^b$ Volker C. Radeloff,$^c$ and David S. Wilcove$^{a, d}$
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: |
  | $^a$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ
  | $^b$Department of Geography, Kent State University, Kent, OH
  | $^c$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI
  | $^d$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ
  |
  | $^*$Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ
output:
  bookdown::word_document2:
    reference_docx: /Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx
    number_sections: false
  bookdown::pdf_document2:
    toc: false
    keep_tex: true
    includes:
      in_header: "preamble.tex"
editor_options:
  chunk_output_type: console
bibliography: [/Users/christophercrawford/work/library/library.bib, packages.bib]
link-citations: yes
csl: conservation-letters.csl
urlcolor: blue
linkcolor: black
citecolor: black
---
```{r misc-notes, include = FALSE, eval = FALSE}
# Authors and Affiliations, formatted to look nice in Word (i.e., nested within the abstract):
# author: Christopher L. Crawford,$^a$$^*$ He Yin,$^b$ Volker C. Radeloff,$^c$ and David S. Wilcove$^{a, d}$
# date: "`r format(Sys.time(), '%B %d, %Y')`"
# abstract: |
#   | $^a$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ
#   | $^b$Department of Geography, Kent State University, Kent, OH
#   | $^c$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI
#   | $^d$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ
#   |
#   | $^*$Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ




# formatted for LaTeX
# author:
#   - Christopher L. Crawford$^1$^[Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ]
#   - He Yin$^2$
#   - Alex Wiebe$^{4}$
#   - Volker C. Radeloff$^3$
#   - David S. Wilcove$^{1,4}$
# date: 
#   $^1$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ\newline
#   $^2$Department of Geography, Kent State University, Kent, OH\newline
#   $^3$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI\newline
#   $^4$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ\newline
#   \newline
#   \today
```

```{r load-packages-functions, eval = TRUE, echo = FALSE, include = FALSE}
source("/Users/christophercrawford/work/projects/biodiversity_abn/scripts/0_start.R")
```

```{r preview-chapter, eval=FALSE, include = FALSE}
my_format <- "simple"
rmarkdown::render("scripts/biodiv_abn_MS.Rmd",
                  output_file = 
                    paste0(p_proj, "writing/previews/", "Crawford_bd_abn_MS_draft",
                           format(Sys.Date(), ("_%Y_%m_%d")), ".docx"))

# Save final PDF with SI for submission:
# my_format <- "latex"
# rmarkdown::render("scripts/Crawford_biodiv_abn_MS+SI.Rmd",
#                   output_file = paste0(p_proj, "writing/", "Crawford_biodiv_abn_MS+SI",
#                            format(Sys.Date(), ("_%Y_%m_%d")), ".pdf"),
#                   output_format = "bookdown::pdf_document2")

# bookdown::preview_chapter(
#   input = "abn_trajectories_ms.Rmd",
#   output_format = "bookdown::word_document2",
#   output_file = paste0("abn_duration_draft", format(Sys.Date(), ("_%Y_%m_%d")), ".docx"),
#   output_dir = paste0(p_proj, "writing/previews"),
#   output_options = 
#     list(reference_docx = "/Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx"),
#   preview = TRUE)
```

```{r set-chunk-opts, include = FALSE}
# set default chunk options
knitr::opts_chunk$set(
  out.width = "1\\textwidth",  # if compiling a pdf, include this option
  # comment = NA,
  # fig.width=6, fig.height=6, 
  echo = FALSE # the code itself will NOT be printed with results
  )

options(knitr.table.format = function() {
  if (knitr::is_latex_output())
    "latex" else "simple"
})
```

```{r produce-package-bib, include = FALSE, echo = FALSE}
# needed_packages
# github_packages
knitr::write_bib(x = c(needed_packages, github_packages), file = 'packages.bib')
# knitr::write_bib(x = c("car"), file = 'extra.bib')

# citation(package = needed_packages[1])

# knitr::write_bib(x = c(needed_packages[c(14)]), file = 'packages_top.bib')
# knitr::write_bib(x = c(needed_packages[3]), file = 'terra.bib')

# 
# needed_packages[1]
# packageDescription(needed_packages[1])
# 
# tinytex::tlmgr_install(
#   # pkgs = "ragged2e"
#   # pkgs = "standalone"
#   pkgs = "preview"
# )
```

```{r load-files, include = FALSE, eval = FALSE}

# set run:
run_label
aoh_run_date

# Additional paths --------------------------------------------------- #
p_plots <- "/Users/christophercrawford/work/projects/abandonment_trajectories/output/plots/"
p_derived2 <- paste0(p_dat_derived, run_label, "/derived_data/")

# load general spatial data on site locations --------------------------------------------------- #
site_df <- read.csv(file = paste0(p_dat_derived, "site_df.csv"))
site_df_w_size <- read.csv(file = paste0(p_dat_derived, "site_df_w_size.csv"))
site_sf <- st_read(paste0(p_dat_derived, "site_sf.shp"))


# load all length data (distilled)
length_distill_df <- read_csv(file = paste0(p_derived2, 
                                            "length_distill_df", run_label, ".csv"))

# load mean length data
mean_length_df <- read_csv(file = paste0(p_derived2, "mean_length_df", run_label, ".csv"))


# load summary statistic derived from lengths
summary_stats_all_sites <- read_csv(file = paste0(p_derived2, "summary_stats_all_sites", run_label, ".csv"))
summary_stats_all_sites_pooled <- read_csv(file = paste0(p_derived2, "summary_stats_all_sites_pooled", run_label, ".csv"))

# load area summary stats:
area_summary_df <- read_csv(file = paste0(p_derived2, "area_summary_df", run_label, ".csv"))
area_recult_threshold <- read_csv(file = paste0(p_derived2, "area_recult_threshold", run_label, ".csv"))


# load stats on proportion of abandoned cropland by land cover
abn_lc_area_2017 <- read_csv(file = paste0(p_derived2, "abn_lc_area_2017", run_label, ".csv"))
abn_prop_lc_2017 <- read_csv(file = paste0(p_derived2, "abn_prop_lc_2017", run_label, ".csv"))


# load combined persistence, area, and turnover datasets:
# also known as dat_l in "4_decay_models.Rmd"

persistence_dat <- 
  # dat_l <- 
  read_csv(file = paste0(p_derived2, "persistence_dat", run_label, ".csv")) %>%
  mutate(site = as.factor(site),
         bins = as.factor(bins),
         year_abn_bins = as.factor(year_abn_bins),
         cohort = as.factor(cohort))

area_dat <- read_csv(file = paste0(p_derived2, "area_dat", run_label, ".csv"))
turnover_dat <- read_csv(file = paste0(p_derived2, "turnover_dat", run_label, ".csv"))



# load decay rate data
load(file = paste0(p_derived2, "/decay_mod_coef_dfs.rds"), 
     verbose = TRUE)

# load extrapolation data
extrapolate_df_l3 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3", run_label, ".csv"))
extrapolate_df_l3_a2 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3_a2", run_label, ".csv"))


# load data for shaanxi maps --------------------------------------------------- #

# load data for decay models --------------------------------------------------- #
fitted_df_l3 <- read_csv(file = paste0(p_derived2, "/decay_mod_fitted_df_l3", run_label, ".csv"))

load(file = paste0(p_derived2, "/decay_mod_coef_dfs.rds"), verbose = TRUE)
# contains:
# coef_df_l3, coef_df_wide_l3, mean_coef_df_l3, mean_coef_df_wide_l3, mean_coef_proj_l3, time_to_l3


# load extrapolation data
extrapolate_df_l3 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3", run_label, ".csv"))

future_df_l3 <- read_csv(file = paste0(p_derived2, "/future_df_l3", run_label, ".csv"))

```

Preparing manuscript for [*Conservation Letters*](https://conbio.onlinelibrary.wiley.com/hub/journal/1755263x/homepage/forauthors.html):

- 3000 word limit
- No more than 6 figures and tables
- 40 References

Structured as follows:  

1. Introduction. The Introduction should summarize briefly the background and aims, and end with a brief statement of what has been achieved by the work.
2. Methods. This section should contain sufficient detail so that all procedures can be clearly understood by the reader. Further technical details can be given in Supporting Information, which should be detailed enough that all procedures can be repeated (in conjunction with cited references).
3. Results. The Results section should present the evidence that supports the conclusions to be drawn in the Discussion. The Results section should conform to a high standard of rigour. Extended lines of inference, arguments or speculations should not be placed in the Results.
4. Discussion. The Discussion section should be separate from the Results section. It allows authors to propose their interpretation of the results, and to suggest what they mean in a wider context. It should end with a clear statement of the main conclusions of the research, and a clear explanation of their importance for conservation policy or practice.
5. Acknowledgements and Data. Acknowledgements of financial or institutional support should be brief (100 words maximum). In this section, authors are also expected to disclose any potential conflicts of interest, financial or otherwise, that a reasonable person could construe as possibly influencing the objectivity of the report. The absence of any statement is presumed to mean that there is nothing to disclose. Authors must provide a data accessibility statement, including a link to the repository they have used to archive their data (see policy on Data Sharing and Accessibility). They should also (where relevant) provide a statement of the ethics review process that was adhered to in gathering the data.
6. References


# Abstract (150 word limit) {.unnumbered}

Background/Question/Methods:

Although agricultural expansion and deforestation continues to be a dominant, and concerning, land use trend in many regions, hundreds of millions of hectares of cropland have also been abandoned globally since 1950 due to shifting demographics, sociopolitical upheaval, environmental changes, and globalization.
These abandoned agricultural lands increasingly represent opportunities for the restoration of ecosystems without imposing significant opportunity costs, with substantial potential environmental benefits in terms of both carbon sequestration and increased habitat availability for wildlife.
However, whether and where abandonment presents an opportunity, or a threat, to biodiversity remains unclear, due to variation in species habitat requirements, threat status, and the spatial distribution of habitat types and abandonment.

Moreover, these environmental benefits depend on the longevity of abandonment, which is poorly known.
Here, we use annual 30-m land cover maps to track changes to species' area of habitat (AOH) from 1987-2017 for 2023 bird and mammal species at eleven sites across four continents.
Our high-resolution time series allows us to track dynamic changes in habitat and determine whether species gain or lose habitat on average through time.

We find that ...
most species gain habitat as a direct result of abandonment.
However, many species' habitat was constrained by the recultivation of abandoned croplands. 
In a scenario in which no recultivation occurred, X more species (Y%) had a positive trend in area of habitat, largely due to gains at our relatively more speciose site in Mato Grosso, Brazil.
There was significant variation in both species richness and the impact of abandonment across our sites, but all but one site showed at least double the number species that gain habitat than those that lose habitat. 

Threatened species were even better off...

However, when we calculated changes to AOH across a broader temporal and spatial scale, the balance of losses outweighed the number of gains. 
By calculating trends in AOH in all pixels that were abandoned, but relative to the land cover in 1987, we found that the trend for most species was negative, despite the abandonment of cropland, largely due to insufficient time following abandonment for species to regain habitat originally lost to make way for cultivation.
Similarly, gains from abandonment did not outweigh losses in habitat elsewhere at each site, leading to overall more losers than winners. 

Concerningly, our analysis of changes to AOH show that many species do not gain habitat overall as a result of abandonment, likely because short-term abandonment lacks sufficient time for habitat to regenerate on abandoned croplands through secondary succession.
About equal numbers of mammal species overall lost habitat as a result of abandonment as gained habitat, but 1.5 times as many bird species lost habitat as gained due to abandonment (though this was less dramatic for threatened bird species, at 1.2x).


abandoned pixels relative to 

We then explore how habitat area changed as a result of cropland abandonment by delving into differences between groups of species, looking at IUCN threat status, habitat preferences, habitat age requirements (early successional vs. mature habitat only)

We find that a significant portion of abandoned croplands are recultivated by the end of our time series, substantially curtailing the potential gains in AOH for different species.
We also find that 

Our land cover maps indicate that 
We show

We also investigate how recent abandonment and subsequent regeneration has affected landscape fragmentation. 

This time series shows that while 

agricultural abandonment , increasing levels of abandonment have raised hopes for recovery in some places.

Agricultural expansion continues to be the do

## Keywords {.unnumbered}

# Introduction {.unnumbered}

Outline:  
1. Intro to abandonment as an issue.
2. Biodiversity is going to be affected by abandonment, but controversy about how. (Play up the controversy.)
3. New data allows us to leverage detailed annual land cover time series data to understand how area of habitat is affect by abandonment.
4. Our goal, questions: understand how abandonment affects biodiversity by investigating trends in area of habitat during our time series, for various subsets of the landscape. 

## set the stage 

Abandonment is taking place around the world. Given that 

Three to four paragraphs to start

With a better understanding of the timing, persistence, and trajectories of abandoned agricultural land, researchers are also better equipped to ask questions about the environmental implications of abandonment, including how biodiversity is likely to be affected by abandonment. To investigate the biodiversity consequences of abandonment, we couple the abandonment trajectories developed in Chapter 2 with models of biodiversity responses to land-use changes.

Despite a large number of studies investigating the recovery of biodiversity in abandoned agricultural land over time (typically species richness and abundance), few studies have used this information to assess the biodiversity implications of abandonment at a landscape scale by connecting maps of abandonment to biodiversity information.
While previous studies have looked at the carbon implications of regeneration in abandoned land [@Nunes2020], the few studies assessing the biodiversity consequences of abandonment have only done so coarsely. 
For example, @Meyfroidt2016 simply use distance to protected areas as a proxy for biodiversity value, and @Henry2019 use only coarse measures of species richness [how was species richness inferred?] and CI biodiversity hotspots. 
In perhaps the best spatial assessment of the potential biodiversity value of abandonment, @Crouzeilles2020 estimate two measures of the biodiversity impact of recent and predicted future forest regeneration in the Atlantic Forest: 1) an intactness measure based on the proportion of each 200 ha pixel that is forested, and 2) the predicted number of species extinctions avoided as a result of habitat gains (based on AOH changes due to regeneration). However, these measures do not consider patterns of persistence or the time-dependent recovery of biodiversity in regeneration habitats as a function of age, nor do they investigate nuance in which species benefit (particularly threatened species). 

To fill these gaps, we calculate direct changes in Area of Habitat (AOH) for terrestrial vertebrate species resulting from abandonment (following @Strassburg2019). 

Note, this paper would be more focused on the breakdown of winners and losers species, trying to answer the question of where abandonment is likely to be a good thing to encourage, or a bad thing to attenuate. 
It would help provide nuance to the perspectives about opportunity or threat in different regions.


## Questions, approach

Here we leverage species range maps and habitat preferences for X species of birds and mammals to calculate changes in area of habitat (AOH; @Brooks2019) resulting from cropland abandonment at 11 sites. 
Our goal is to understand how abandonment affects biodiversity by investigating trends in area of habitat during our time series, for various subsets of the landscape. 
To what extent did cropland abandonment result in gains, or losses of habitat across the bird and mammals species that occur at each of our sites?

We address three questions:  

1. Does abandonment create more habitat, or less, for bird and mammal species, and how does the proportion of winners and losers vary across sites and across different groups of species, including by IUCN threat status, and by whether species are mature forest obligates or not? (Not emphasizing range size...) 
2. Reculivation, by comparing observed vs. potential maps of abandonment, and habitats.
3. How do gains resulting from abandonment compare to changes in habitat over longer time scales, or larger geographic scopes?

(Summary statement)
Together, our analysis indicates that cropland abandonment likely benefitted a plurality of bird and mammals species, but trends in area of habitat varied widely. 
More species gained habitat than lost as a result of cropland abandonment, but rarely compensates for habitat loss that occurred either earlier in the time series to make way for cropland, or that occurred elsewhere in the site (irrespective of abandonment).
The share of species that gained habitat on net stood to increase slightly in our scenario without recultivation, but on the whole this did not amount to a huge change. (How much did species AOH in 2017 increase between observed and potential?)

as evidenced by



# Methods {.unnumbered}

Outline:

Look to "short_methods"



## Abandonment maps and definitions

We build on a 30-m annual land cover time series produced for eleven sites from 1987-2017 [@Yin2020].
These sites, each between 1.6 and 5.7 million ha (Mha) in size, were selected to represent areas with documented histories of cropland abandonment in the past three decades, and cover a range of biomes, including drylands, temperate regions, and the sub-tropic and wet tropics, and span four continents.
(We excluded three sites from those originally mapped by @Yin2020 that had low classification accuracy.)
This time series mapped four land cover classes at high accuracy (average overall accuracy of $85\pm4$%): 1) cropland, 2) herbaceous vegetation (e.g., grassland), 3) woody vegetation (e.g., forests), and 4) non-vegetation (e.g., water, urban, or barren land).
Previous work [@Crawford2022] mapped abandoned croplands by identifying periods of non-cropland (grassland or woody vegetation) that followed periods of cultivation and remained non-cropland for five or more consecutive years, following a commonly used five-year abandonment definition designed to exclude short-term fallow cycles [@FAO2016].
We used these abandonment maps to identify all pixels that were abandoned at least once during the time series (the "maximum extent of abandonment").

In order to assess the implications of recultivation on the impact of abandonment on biodiversity, we use a simple scenario of "potential abandonment" developed by [@Crawford2022] that assumed no recultivation took place during the time series. 
In this scenario, abandoned croplands were left abandoned from the year of initial abandonment through the end of the time series.

## Translating land cover to IUCN habitat types

Because AOH is calculated based on habitat, not land cover, we translated our land cover maps into maps of habitats to match the habitat suitabilities listed by IUCN. 
To do so, we used a recently published 100-m global map of IUCN habitat types for the year 2015 developed by @Jung2020a, resampled to 30-m to match our land cover maps.
First, we developed a simple crosswalk between our coarse land cover types (woody vegetation, herbaceous vegetation, non-vegetation, and cropland) and the more specific habitat types considered by the IUCN (e.g., temperate forest, subtropical/tropical dry grassland), based on how each habitat type would likely be categorized during classification of satellite imagery into land cover (see Table \@ref(tab:iucn-crosswalk-table)).
Because this crosswalk delineates "one to many" relationships (i.e., one land cover class could be multiple IUCN habitat types), we next assigned individual pixels of land cover to the most common corresponding IUCN habitat type within the surrounding pixels in the 2015 habitat map from @Jung2020a (i.e., assigning a coarse land cover like grassland to the most common grassland habitat type within the surrounding neighborhood; see Section \@ref(si-habitat-assignment) for more details).
We considered shrubland and savanna habitat types to be a subset of our woody vegetation (i.e., forest) and herbaceous vegetation (i.e., grassland) land cover classes, respectively.
Because these habitats occupy intermediate zone between these land cover classes, we conducted a sensitivity analysis to assess similarity between land cover and reclassified habitat maps under the full range of possible categorizations (see Section \@ref(si-savanna-shrubland-jac) and Fig. \@ref(fig:jung-crosswalk-jac). 
We found similar levels of spatial overlap across all possible categorizations, so we categorized these habitats to match how they would be classified directly from satellite imagery (as noted previously).

## Species range maps and habitat preferences

Our analysis is built on the calculation of a species' Area of Habitat (AOH, *sensu* @Brooks2019), defined as the total area of a species range within a given site that falls within suitable habitat types and elevations, as derived from IUCN assessments.
We sourced species range maps for all bird and mammal species from Birdlife International and the IUCN respectively, and extracted habitat and elevational preferences from IUCN assessments. 
We exclude marine species and species classified as “Extinct” (EX) or “Extinct in the Wild” (EW), and only considered areas where species were considered “extant” (presence) and “native” or “reintroduced” (origin). 
For species with range maps delineated by season, we only consider areas where species are “resident,” “breeding,” or “non-breeding” (we exclude “passage” areas for migratory species).
For species with range maps delineated by season (i.e., "resident," "breeding," and "wintering" ranges), we only considered those habitats marked as "suitable" in the relevant season associated with that portion of the range.

When comparing results by IUCN Red List status, we broadly defined species as "threatened" when assessed by the IUCN as "Vulnerable," "Endangered," or "Critically Endangered," and "not threatened" when assessed as "Near Threatened" or "Least Concern."
Species assessed as "Data Deficient" are excluded.
We classified species by range size based on whether a species total global range area (extent of occurrence) was above or below the global median (following @Jenkins2013).

Because species' habitat preferences listed in IUCN assessments do not account for habitat age, we classified species based on whether they required mature habitat in at least one season at each site, here defined as any habitat requiring at least 30 years of secondary succession to become suitable for that species (the maximum age of abandoned croplands was 30 years, given the bounds of our time series).
Species were manually classified into mature forest obligates and non-obligates based on habitat descriptions from IUCN assessments and individual species accounts in The Handbook of the Birds of the World [@BOW2022] and The Handbook of the Mammals of the World [@HMW2019].^[*Note: Chris has marking species that are on the border as either 0.1, 0.25, 0.75, or 0.9, based on level of confidence. Mammals have not yet all been classified.*] 
Neotropical bird species were independently verified by expert observer R. Alex Wiebe.
These species are conservatively excluded throughout, though patterns do not change substantially.
(If anything, patterns of AOH trends for abandoned pixels during the full time series, as well as for the full landscape are likely to be even more negative when including these species, given that they are unlikely to benefit from abandonment on the time scale we observe.)

## Calculating Area of Habitat at different temporal and spatial scales

We calculated AOH across three geographic and temporal scales: 1) abandonment only, starting from first cultivation, 2) abandonment only, relative to 1987, and 3) full landscape, 1987-2017.
We first considered changes in habitat stemming directly from cropland abandonment. 
This involved calculated AOH across only pixels that were abandoned at some point during the time series, derived by constraining our time series of habitat types (described above) by the maximum extent of abandonment taken from @Crawford2022.
Furthermore, in order to isolate the effect of cropland abandonment on species' habitat, we only considered habitats in abandoned pixels from the year a given pixel was first cultivated through the end of the time series (i.e., "from first cultivation through abandonment").
Second, we considered changes in habitat that took place in abandoned pixels relative to the start of our time series.
Like our first AOH calculation, we restricted AOH to only pixels that experienced abandonment, but now considered the complete trend in habitat across the full time series (1987-2017), in order to capture any habitat changes that occurred before cultivation and subsequent abandonment (e.g., deforestation).
Third, we calculated AOH for the full landscape at each site, across the full time series (1987-2017), therefore capturing the overall trend in AOH that each species experienced during out study period, including abandonment but also other land cover change dynamics. 

@Crawford2022 showed that while substantial abandonment occurred at our eleven study sites (8.76 Mha of croplands were abandoned one or more times from 1987-2017), this abandonment was relatively short-lived and subject to high rates of recultivation (only 6.06 Mha of croplands remained abandoned as of 2017).
In order to understand how the relative ephemerality of abandonment affected the extent to which species gained or lost habitat as a result of abandonment, we also calculated AOH for an estimate of potential abandonment, which explored a scenario in which no recultivation took place following abandonment (i.e., all abandoned pixels remained abandoned from the year of initial abandonment through the end of the time series, following @Crawford2022).
AOH for potential abandonment was calculated similarly to our first  was 

## Summarizing AOH trends

We extracted trends in AOH over time for each species at each site using simple linear regressions parameterized to predict AOH as a function of year.
AOH trends were classified as either gains or losses of habitat over time when slope coefficients were positive or negative, respectively, and statistically significant (p < 0.05, where the 95% confidence interval of the slope coefficient did not overlap with zero).
Species with AOH slopes that were not statistically significant (p >= 0.05) were classified as having "No Trend" at a given site. 
Because each species could occur at multiple sites, there were some cases where a individual species had multiple types of trends at different sites.
For overall results counting species' trends across all sites, we grouped species based on the consistency of trend direction.
Species with consistently positive trends ("Gain"), negative trends ("Loss"), or insignificant trends ("No Trend") across all sites where they occur were classified based on the single trend. Species with a significant trend (gain or loss) at one or more sites, but no significant trend at one or more additional sites, were considered to have a "weak" trend overall ("Weak Gain" or "Weak Loss").
Species with significant gains at one or more sites, and significant losses at others, were considered to have "Opposite" trends, where the impact of abandonment is context dependent.

Trends were subsequently summarized across various groups of species, including by site, by IUCN threat status, by range size, and whether a species was considered a mature forest obligate species.




## Data processing

We analyzed spatial data (maps of land cover, abandonment, IUCN habitat types, species ranges, and elevation), IUCN assessment data, and resulting area of habitat data in RStudio version `r rstudioapi::versionInfo()$version` [-@RStudio], using `r R.Version()$version.string` [-@R-base], using the terra [@R-terra], data.table [@R-data.table], and tidyverse [@R-tidyverse] packages.

# Results {.unnumbered}

We address three questions:

1. Does abandonment create more habitat, or less, for bird and mammal species, and how does the proportion of winners and losers vary across sites and across different groups of species, including by IUCN threat status, and by whether species are mature forest obligates or not? (Not emphasizing range size...) 
2. Reculivation, by comparing observed vs. potential maps of abandonment, and habitats.
3. How do gains resulting from abandonment compare to changes in habitat over longer time scales, or larger geographic scopes?

## Outline

The results are structured as follows:
First, results overall, for just cropland abandonment. I have the first overall figure, but maybe I'd be better off starting with the site-by-site figure. 

Mention birds vs. mammals, then go into (all for crop_abn)
- threatened species
- small-ranged species
- mature forest obligates. 

Then, present observed vs. potential abandonment. 

Finally, discuss cropland abandonment in the context of the full time series, and the full landscape. 

Put the maps at the end, I think. 

New figure idea:
- new summary figure, combining the summary trend (just for cropland abandonment) and site by site results
- perhaps a single bar, showing the relative proportion. 
- Or, just the summary on the left side of a figure, and the site by site numbers on the right
- Threatened species by site. (Maybe I can add this with shading)

######## 

1. Overall
- By site
- Birds vs. Mammals
- By IUCN threat status
- Range size (simple mention)
- Mature forest obligates

2. Observed vs. potential abandonment:

Does recultivation
Two results here: a) how the share of winners vs. losers changes between the two scenarios, and b) how the AOH would change, in terms of percent increase in habitat, on average across species. 

3. Comparing trends when calculating AOH across longer time scales and large geographic scopes (AOH Types)

Cropland abandonment *per se* compared with trends in abandoned pixels for 1987-2017, and compared to the full landscape (1987-2017).
Our initial results on cropland abandonment *per se* tell us whether abandonment of cropland causes a gain or loss of habitat for each species. 
By looking at the trend for those same abandoned pixels, but starting from the beginning of our time series (1987), puts gains from abandonment in the context of any habitat loss that occurred prior to cultivation.
Taking this one step further, by assessing trends of AOH for the full site area across the full time series, we put our results in context of broader trends taking place at each site. 


Overall summary:

Together, our analysis indicates that cropland abandonment likely benefitted a plurality of bird and mammals species, but trends in area of habitat varied widely. 
More species gained habitat than lost as a result of cropland abandonment, but rarely compensates for habitat loss that occurred either earlier in the time series to make way for cropland, or that occurred elsewhere in the site (irrespective of abandonment).
The share of species that gained habitat on net stood to increase slightly in our scenario without recultivation, but on the whole this did not amount to a huge change. (How much did species AOH in 2017 increase between observed and potential?)


Thoughts from figures:

2. Overall, side by side.
(Maybe this should be moved later in the paper?)
For cropland abandonment alone, the majority of bird and mammal species gain habitat (i.e., have positive trends in AOH over time).
This is a very clear trend - there are ~700 bird species with statistically significant positive trends in AOH over time, and ~400 mammal species, across all sites, compared to about ~125 bird and ~15-20 mammal species that lose habitat as a direct result of abandonment. 

(Maybe a separate point, or moved to a later section about the different scales of AOH.)
This pattern is reversed for birds, however, when considering the full time series, even when only considering abandoned pixels.
This is likely because many species might lose habitat during conversion to cropland, but not gain it entirely back as a result of abandonment, whether due to recultivation or a lack of time for succession. 
This pattern is even stronger for mature forest obligate bird species (Fig. \@ref(fig:aoh-overall-mature-for-obl)). 
Even when excluding these species, the balance of winners to losers still shows more losers than winners for both abandonment (1987-2017) and the full landscape (1987-2017).

3. By site
All sites showed a greater number of bird and mammal species that gained habitat from abandonment, with a single exception of birds in Iraq (which is in the most arid biome, and with the lowest site species richness (Fig. \@ref(fig:richness-by-site)).
The relative proportion of winner to loser species varied across sites, but at most sites, at least twice as many species gained habitat as lost habitat, and often more. 
On the whole, a greater proportion of bird species lost habitat as a result of abandonment, compared with mammals (the proportion of winners to losers was about the same for Mato Grosso; Fig. \@ref(fig:aoh-by-site-stacked).

4. Trend lines, by site
The trend lines (Fig. \@ref(fig:aoh-trend-lines-crop-abn)) showed that some sites had relatively steeper trends, on average.
Not sure if this is a main text figure.

5. By IUCN Status.

A relatively higher proportion of threatened species gained habitat as a result of abandonment, though there were relatively low numbers of threatened species at each site (Fig. \@ref(fig:richness-by-site)).

species were relatively (see also Fig. \@ref(fig:aoh-by-iucn) for all IUCN statuses, and Figs. \@ref(fig:aoh-by-site-threatened-crop-abn)-\@ref(fig:aoh-by-site-threatened-full) for counts of species in each trend for threatened species only.)

6. Obs. vs. potential - move earlier. 

If no recultivation had occurred during our time series (i.e. our scenario of "potential" abandonment), the share of species that gain habitat would have increased for both birds and mammals.
Moreover, many species with a "Weak Gain" trend (indicating gains at some sites, but no trend at others), would experience consistent gains across sites in our potential abandonment scenario.
However, a greater number of species had statistically significant losses of habitat in this scenario of potential abandonment, likely due to species with no trend or opposite trends converting into consistent losing trends.

At most sites, the differences in proportions between observed and potential abandonment were not large.
However, the single exception was Mato Grosso, Brazil, where 70 additional bird species showed statistically significant positive trends in AOH during our time series (231 compared to 161), largely resulting from the transition of many species from "no trend" to "gain" species (excluding mature forest obligate birds; Fig. \@ref(fig:aoh-obs-v-pot-by-site-bird)).
This is a significant number of bird species!

```{r}
aoh_trends %>%
  filter(passage_type == "exclude_passage",
         vert_class == "bird", grepl("crop", aoh_type),
         mature_forest_obl < 0.5) %>%
  group_by(aoh_type, vert_class, passage_type, site, trend) %>%
  summarise(n_sp = n()) %>% 
  filter(site == "mato_grosso") %>%
  arrange(trend)
```



(ref:caption-aoh-by-site-crop-abn) Number of bird and mammal species shown by trend in AOH in abandoned pixels starting from the year of first cultivation and running through 2017. Species counts are shown for individual sites and summed across all sites ("Overall" at left).

```{r aoh-by-site-crop-abn, fig.cap = '(ref:caption-aoh-by-site-crop-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_overall_w_site_combo_crop_abn_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-trend-lines-crop-abn) AOH trends for all bird and mammals species occurring at each site. AOH is calculated for only pixels that were abandoned at least once, from the year of first cultivation through 2017.

```{r aoh-trend-lines-crop-abn, fig.cap = '(ref:caption-aoh-trend-lines-crop-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_crop_abn_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-iucn-lumped) Proportion of species with various trends in AOH, shown for species classified by the IUCN as "threatened" ("Vulnerable," "Endangered," or "Critically Endangered") or "not threatened" ("Near Threatened" or "Least Concern"). Species classified as "Data Deficient" are excluded.

```{r aoh-by-iucn-lumped, fig.cap = '(ref:caption-aoh-by-iucn-lumped)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_IUCN_status_lumped",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-obs-v-pot) Proportion of species that gain or lose AOH as a result of abandonment during our time series, for both observed abandonment and a scenario of potential abandonment, assuming no recultivation.

```{r aoh-obs-v-pot, fig.cap = '(ref:caption-aoh-obs-v-pot)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_obs_v_pot_overall",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-overall) Number of bird and mammal species with various trends in AOH during our time series. Columns correspond to three scales of AOH calculation: only abandoned pixels, from first cultivation through 2017; only abandoned pixels, from 1987-2017; and the full landscale at each site, from 1987-2017. Trends in AOH were determined through linear regression of each species AOH over time at each site. Trends that were not statistically significant (p >= 0.05) are shown as "No Trend." Because each species could occur at multiple sites, there were some cases where a individual species had multiple types of trends at different sites. Species with consistently positive trends ("Gain"), negative trends ("Loss"), or insignificant trends ("No Trend") are shown in dark green, yellow, and black, respectively. Species that had a significant trend at one or more sites, and no trend at one or more other sites, were considered to have a Trend of "Weak Gain" or "Weak Loss." Species with significant gains at one or more sites, and significant losses at others, are marked as "Opposites" (where the impact of abandonment is context dependent). (NB: Chris will remove mature forest obligate bird species from this, though see \@ref(fig:aoh-overall-mature-for-obl)) 

```{r aoh-overall, fig.cap = '(ref:caption-aoh-overall)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_by_vert_aoh_type", aoh_run_date,".pdf"))
```

(ref:caption-abn-lc-2017) Land cover as of the year 2017, shown for all pixels that were abandoned at least once during the time series.

```{r abn-lc-2017, fig.cap = '(ref:caption-abn-lc-2017)'}
knitr::include_graphics(path = paste0(p_plots, "abn_land_cover_2017.pdf"))
```

# Discussion {.unnumbered}

## Study limitations {.unnumbered}

Outline:

- Fragmentation
- Habitat preferences as a function of age of habitat, succession. Regeneration can be successful, but it typically takes a long time (> 50 years) for regenerating habitats to recover levels of species richness, and even longer to recover to similar community compositions.
- Study site size: because our land cover time series was restricted to footprint of our sites, 

## Policy implications {.unnumbered}



# Acknowledgements {.unnumbered}

We thank the following researchers at the University of Wisconsin-Madison, who generously shared the land cover maps that made our analysis possible: Amintas Brandão, Johanna Buchner, David Helmers, Benjamin G. Iuliano, Niwaeli E. Kimambo, Katarzyna E. Lewińska, Elena Razenkova, Afag Rizayeva, Natalia Rogova, Seth A. Spawn-Lee, and Yanhua Xie.
We thank the Drongos research group for advice and companionship.
Our analysis would not be possible without the dedication of IUCN, BirdLife International, and numerous individual experts who contributed to species assessments.
Analyses were performed using Princeton Research Computing resources at Princeton University.
This research was supported by the High Meadows Foundation (DSW) and the NASA Land Cover and Land Use Change Program (Grant no. 80NSSC18K0343, to VCR, HY), and contributes to the Global Land Programme (<https://glp.earth/>).

## Author contributions {.unnumbered}

TBD

<!-- CLC and DSW conceived of the project in collaboration with HY and VCR. -->
<!-- HY contributed guidance on data curation and the analysis of raw land cover data. -->
<!-- CLC designed and conducted the data analysis, with feedback from HY, DSW, and VCR. -->
<!-- All authors contributed to writing the manuscript, with CLC preparing the initial draft. -->

## Competing interests {.unnumbered}

The authors declare no competing interests.

## Data and code availability {.unnumbered}

The annual land cover maps that underlie our analysis are archived at Zenodo (<https://doi.org/10.5281/zenodo.5348287>), along with other datasets generated and analyzed for this article (including maps of abandonment age and datasets summarizing area, persistence, and turnover). 
Code to replicate these analyses is available on GitHub at <https://github.com/chriscra/biodiversity_abandonment>, and is archived at Zenodo (<TBD>).
Bird and mammal range maps are available upon request from BirdLife International (<http://datazone.birdlife.org/species/requestdis>) and IUCN (<https://www.iucnredlist.org/resources/spatial-data-download>) respectively.
Species assessment data (including habitat and elevation preferences) are available from IUCN (<https://www.iucnredlist.org/>) 
The 2015 global map of IUCN habitat types developed by @Jung2020a is freely available at <https://zenodo.org/record/4058819>.



# References {.unnumbered}

BirdLife International and Handbook of the Birds of the World (2021) Bird species distribution maps of the world. Version 2021.1. Available at http://datazone.birdlife.org/species/requestdis.

Request blurb:
This IUCN Assessment data will be used for a dissertation research project designed to investigate the impact of agricultural abandonment on biodiversity. The final product will be academic journal articles, which will be shared with the IUCN science team upon completion. Please don't hesitate to get in touch if you have any questions (ccrawford@princeton.edu). Thank you! - Christopher Crawford, PhD Candidate, Princeton University.

Range maps downloaded on January 4th, 2022 (amp, mam, reptiles, from IUCN, and birds from BirdLife International)

::: {#refs}
:::

\newpage

\beginsupplement

# Supplementary information {.unnumbered}

\tableofcontents
\listoftables
\listoffigures

\newpage

# Extended Results {#results-si}

# Extended Methods

## Assigning land cover to habitat types {#si-habitat-assignment}

We developed a method for assigning pixels in our coarse land cover classes to finer habitat types to match the habitats used by the IUCN to describe species' habitat preferences.
This process was based on a 2015 map of IUCN habitat types at 100-m resolution produced by @Jung2020a based on a dynamic crosswalk between the Copernicus land cover maps and the IUCN habitat types.
We in turn developed a simple crosswalk between our coarse land cover types and the finer IUCN habitat types mapped by @Jung2020a, as shown in Table \@ref(tab:iucn-crosswalk-table). 
But because there was not perfect spatial overlap between our land cover maps and the 2015 habitat map, we developed a method for assigning our pixels of land cover to the nearest habitat type on the @Jung2020a map that matched the broad category according to our crosswalk. 
We first extracted only those pixels with habitat types that fell into each of our four land cover classes, and used a focal filter to assign habitat types to empty pixels based on the most common (i.e., modal) habitat type within the surrounding nine-pixel neighborhood (using the function `terra::focal(fun = "modal")`).
This produced four maps of potential habitat types for each pixel corresponding to each coarse land cover type.
We then masked each of these potential habitat types maps by the corresponding land cover pixels in each year at each site, producing maps of the likely habitat type for each pixel of land cover, in each year.
We then combined these four back together again to produce a map of habitat types at each site in each year, which served as input for our calculation of area of habitat (AOH). 

See maps below.

### Allocating savanna and shrubland habitats to land cover classes {#si-savanna-shrubland-jac}

Allocating the finer scale IUCN habitat types to our coarse land cover classes was straightforward for forests, grasslands, non-vegetation habitat types, and arable land. 
It was less straightforward for savanna and shrubland habitats, which are intermediate between entirely open grasslands and closed forests.


Sites with significant area classified as shrubland include:
- Belarus (14%)
- Bosnia & Herzegovina (7%)
- Wisconsin (9%)
- Iraq, Nebraska, and Shaanxi (1-2%)
- Chongqing, Volgograd, and Orenburg (~0.5%)

Sites with significant area in savanna include:
- Goias (69%)
- Chongqing (19%)
- Mato Grosso (10%)

Our default was to classify savanna and shrubland habitats as grassland and forest land cover respectively, but we explored the consequences of alternative classifications by testing spatial similarity between the resulting reclassified maps. 

We calculated Jaccard similarity between our @Yin2020 land cover maps (2015) and @Jung2020a habitat maps (2015) when reclassifying savanna and shrubland habitats as either forest or grassland.
The Jaccard similarity index is a measure of spatial similarity (i.e., overlap) between two sets, defined as the proportion of shared elements, or the intersection divided by the union (Equation \@ref(eq:jaccard-equation)) [@Legendre2012]:

\begin{equation}
J(a,b) = \frac{a\cap b}{a\cup b} (\#eq:jaccard-equation)
\end{equation}

We tested the full range of reclassifications, and the Jaccard similarity between the reclassified maps are shown in Fig. \@ref(fig:jung-crosswalk-jac).

We found that:

1. At the three sites with significant amounts of shrublands (Belarus, Bosnia & Herzegovina, and Wisconsin) - which are all in forested biomes - the similarity between grassland maps increases when shrublands are considered grasslands (as do forest maps). The gold and black bars look basically the same for these three sites, which means that the increase is due to the shrub -> grassland crosswalk; the crosswalk for savanna does not seem to matter, likely because there is very little savanna at these sites. This pattern also holds for Iraq, which had about 2% shrubland (compared to 49% desert, 42% cropland, 4% grassland, and 2% urban).

2. Sites with significant area in savanna (Goias, Mato Grosso, and Chongqing, though I know this isn’t really “savanna”) show greater similarity when savanna is classed as grassland, not as forest (this makes sense, given that these sites likely have a greater distinction between true forest and savanna). The green and gold bars are the same, signifying that the shrubland decision doesn’t matter much (because there is so little shrubland there).

3. The four remaining sites - Nebraska, Orenburg, Shaanxi, and Volgograd - are all in grassland/steppe biomes, where the majority of the site is classified as grassland. At these places, the similarity is higher when shrubland is classified as forest (shown by the green and gray bars being higher).  These increases are pretty small, however, and there is relatively little shrubland at any of these sites (max. 1.4% at Nebraska)

My conclusions from this are that, across the board, I should classify savanna as grassland. For shrubland, however, because there is variation in how similar land cover maps are under these different reclassification rules, but no clear trend, we consider shrubland habitats to be a subset of forest land cover, given that shrublands are most similar spectrally to forests, from a remote sensing perspective. 

(ref:caption-jung-crosswalk-jac) Jaccard Similarity between @Yin2020 land cover maps (2015) and @Jung2020a habitat maps (2015) when reclassifying savanna and shrubland habitats as either forest or grassland.

```{r jung-crosswalk-jac, fig.cap = '(ref:caption-jung-crosswalk-jac)'}
knitr::include_graphics(path = paste0(p_plots, "crosswalk_jaccard.pdf"))
```

# Supplemental Tables

```{r iucn-crosswalk-table, echo = FALSE}
iucn_crosswalk %>%
  left_join(tibble(lc = 1:4, 
                   lc_name = c("Non-vegetation", "Woody vegetation (Forest)",
                               "Cropland", "Herbaceous vegetation (Grassland)"))
            ) %>%
  select(lc, lc_name, IUCNLevel, map_code) %>%
  arrange(lc) %>% #print(n = 50)
  
  
  knitr::kable(., 
             col.names = c("Land cover code", 
                           "Land cover class", 
                           "IUCN habitat",
                           "IUCN map code"),
             align = "r", 
             #format = "simple",
             format = my_format,
             booktabs = TRUE,
             caption = "Observed area abandoned (Mha) as of 2017, potential area abandoned by 2017 (Mha) assuming no recultivation, the percent difference between observed and potential abandonment, and observed abandonment as of 2017 as a percent of the total cropland extent (i.e., the area of all lands that were cultivated at some point during the time series). This table replicates Figure S2. Note that sites are shown in descending order of area abandoned as of 2017 as a proportion of total cropland extent."
             ) %>%
  kableExtra::kable_styling(
    # latex_options = "scale_down"
    latex_options = "hold_position"
    ) %>%
  # row_spec(11, hline_after = TRUE) %>%
  # row_spec(12, bold = TRUE) %>%
  column_spec(1, width = "1.5cm") %>%
  column_spec(2, width = "3cm") %>%
  column_spec(3, width = "4.1cm") %>%
  column_spec(4, width = "1.5cm")
```

# Supplemental Figures

(ref:caption-richness-by-site) Species richness at each site, or the count of species with some Area of Habitat (AOH) at each site. Richness is shown for threatened ("Vulnerable," "Endangered," or "Critically Endangered") or non-threatened ("Least Concern", "Near Threatened"), as assessed by the IUCN. Species classified as "Data Deficient" are excluded.

```{r richness-by-site, fig.cap = '(ref:caption-richness-by-site)'}
knitr::include_graphics(path = paste0(
  p_plots, "richness_by_site_w_category_facet_color.pdf"))
```

## Maps

Include:
- 

## Area line charts:

Two types of area

Three AOH types:

## AOH Results

### Overall

(ref:caption-aoh-overall-stacked) Overall proportion of trend types across sites, for birds and mammals, for three AOH types (cropland abandonment only, abandonment 1987-2017, and full landscape 1987-2017).

```{r aoh-overall-stacked, fig.cap = '(ref:caption-aoh-overall-stacked)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_stacked", aoh_run_date,".pdf"))
```

(ref:caption-aoh-overall-mature-for-obl) Number of species with various trend types across all sites, shown for mature forest obligate and non-obligate birds.

```{r aoh-overall-mature-for-obl, fig.cap = '(ref:caption-aoh-overall-mature-for-obl)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_overall_by_mature_obl",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-overall-mature-for-obl-stacked) Proportion of trend types across all sites, shown for mature forest obligate and non-obligate birds.

```{r aoh-overall-mature-for-obl-stacked, fig.cap = '(ref:caption-aoh-overall-mature-for-obl-stacked)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_overall_stacked_by_mature_obl",
                                      aoh_run_date,".pdf"))
```



### By site

(ref:caption-aoh-by-site-full) Number of species that gain or lose AOH over time across sites, where AOH is calculated for full site area, from 1987-2017 ("full_iucn").

```{r aoh-by-site-full, fig.cap = '(ref:caption-aoh-by-site-full)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_overall_w_site_combo_full_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-max-abn) Number of species that gain or lose AOH over time across sites, where AOH is calculated in only those pixels that were abandoned during the time series, calculated for 1987-2017 ("max_abn_iucn").

```{r aoh-by-site-max-abn, fig.cap = '(ref:caption-aoh-by-site-max-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_overall_w_site_combo_max_abn_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-stacked) Proportion of species at each site with trends in AOH, shown across three scales of AOH calculation.

```{r aoh-by-site-stacked, fig.cap = '(ref:caption-aoh-by-site-stacked)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_site_stacked_3",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-mature-for-obl) Proportion of species with various trends in AOH at each site, shown for mature forest obligate and non-obligate birds, across three scales of AOH calculation.

```{r aoh-by-site-mature-for-obl, fig.cap = '(ref:caption-aoh-by-site-mature-for-obl)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_site_stacked_no_mat_obl",
                                      aoh_run_date,".pdf"))
```

### By IUCN

(ref:caption-aoh-by-iucn) AOH trends, in terms of the proportion of species with a given trend direction, shown by IUCN Red List Status.

```{r aoh-by-iucn, fig.cap = '(ref:caption-aoh-by-iucn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_IUCN_status",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-threatened-crop-abn) AOH trends shown at each site for only those species classified as "threatened" by the IUCN ("Vulnerable," "Endangered," or "Critically Endangered"). Shown for only pixels that were abandoned at least once during the time series, from cultivation through the end of the time series.

```{r aoh-by-site-threatened-crop-abn, fig.cap = '(ref:caption-aoh-by-site-threatened-crop-abn)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/trend_summary_by_IUCN_status_by_site_crop_abn_iucn_Threatened",
  aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-threatened-max-abn) AOH trends shown at each site for only those species classified as "threatened" by the IUCN ("Vulnerable," "Endangered," or "Critically Endangered"). Shown for only pixels that were abandoned at least once during the time series, from 1987-2017.

```{r aoh-by-site-threatened-max-abn, fig.cap = '(ref:caption-aoh-by-site-threatened-max-abn)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/trend_summary_by_IUCN_status_by_site_max_abn_iucn_Threatened",
  aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-threatened-full) AOH trends shown at each site for only those species classified as "threatened" by the IUCN ("Vulnerable," "Endangered," or "Critically Endangered"). Shown for AOH calculated across the full landscape at each site, from 1987-2017

```{r aoh-by-site-threatened-full, fig.cap = '(ref:caption-aoh-by-site-threatened-full)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/trend_summary_by_IUCN_status_by_site_full_iucn_Threatened",
  aoh_run_date,".pdf"))
```

### Observed vs. Potential Abandonment

(ref:caption-aoh-obs-v-pot-by-site-bird) Number of bird species with various trends in AOH for scenarios of observed vs. potential abandonment, shown for each site. Mature forest obligate birds are excluded.

```{r aoh-obs-v-pot-by-site-bird, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site-bird)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_obs_v_pot_by_site_sbs_bird",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-obs-v-pot-by-site-mam) Number of mammals species with various trends in AOH for scenarios of observed vs. potential abandonment, shown for each site.

```{r aoh-obs-v-pot-by-site-mam, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site-mam)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_obs_v_pot_by_site_sbs_mam",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-obs-v-pot-by-site-stacked) Proportion of species with various AOH trends for scenarios of observed vs. potential abandonment, shown for each site

```{r aoh-obs-v-pot-by-site-stacked, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site-stacked)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_obs_v_pot_by_site",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-obs-v-pot-by-site-mature-for-obl) AOH for observed vs. potential abandonment, by site and AOH scale, shown for mature forest obligate and non-obligate birds.

```{r aoh-obs-v-pot-by-site-mature-for-obl, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site-mature-for-obl)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_obs_v_pot_by_site_mat_for_obl",
                                      aoh_run_date,".pdf"))
```

### Range size

(ref:caption-aoh-by-range) AOH by range size.

```{r aoh-by-range, fig.cap = '(ref:caption-aoh-by-range)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_range_side",
                                      aoh_run_date,".pdf"))
```

### Trend Lines:

(ref:caption-aoh-trend-lines-crop-abn-potential) AOH trends for all species occurring at each site, for only abandoned pixels, from cultivation through 2017, for a scenario of potential abandonment.

```{r aoh-trend-lines-crop-abn-potential, fig.cap = '(ref:caption-aoh-trend-lines-crop-abn-potential)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_crop_abn_potential_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-trend-lines-max-abn) AOH trends for all species occurring at each site, for only abandoned pixels, from 1987-2017.

```{r aoh-trend-lines-max-abn, fig.cap = '(ref:caption-aoh-trend-lines-max-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_max_abn_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-trend-lines-full) AOH trends for all species occurring at each site, for the full landscape at each site, 1987-2017.

```{r aoh-trend-lines-full, fig.cap = '(ref:caption-aoh-trend-lines-full)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_full_iucn",
                                      aoh_run_date,".pdf"))
```

### Including Passage Areas

TBD

## Bonus Figures (TBD)

1. (TBD) A summary plot showing Winners/Losers bar chart across different categories, mature forest obligates, range size.
2. (TBD) Results by season (resident, breeding, wintering) How many species have diverging trends across habitats? Perhaps some may be winners in breeding habitat, but losers for wintering habitat.
3. (TBD) Trend of overall habitat vs. AOH in abandonment periods only () vs. AOH in non-abandoned land.
  a. Might be possible to do a linear regression on this middle line. Should this be based on only abandonment periods? This is actually more complicated… it needs to be the complete inverse of “crop_to_abn” maps. In other words, all pixels that were never abandoned plus the noncrop periods that preceded cultivation. This should be compared with the AOH in abandoned pixels only, starting with first cultivation.
4. (TBD) Winners and losers by functional group.
5. (TBD) Mapping the habitat for one species in particular, at one site, with the habitat (maybe future)

