---
title: "Cropland abandonment provides habitat for most species, but rarely compensates for habitat loss"
author: Christopher L. Crawford,$^a$$^*$, R. Alex Wiebe,$^{d}$ He Yin,$^b$ Volker C. Radeloff,$^c$ and David S. Wilcove$^{a, d}$
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: |
  | $^a$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ
  | $^b$Department of Geography, Kent State University, Kent, OH
  | $^c$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI
  | $^d$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ
  |
  | $^*$Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ
output:
  bookdown::word_document2:
    reference_docx: /Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx
    number_sections: false
  bookdown::pdf_document2:
    toc: false
    keep_tex: true
    includes:
      in_header: "preamble.tex"
editor_options:
  chunk_output_type: console
bibliography: [/Users/christophercrawford/work/library/library.bib, packages.bib]
link-citations: yes
csl: conservation-letters.csl
urlcolor: blue
linkcolor: black
citecolor: black
---
```{r misc-notes, include = FALSE, eval = FALSE}
# Authors and Affiliations, formatted to look nice in Word (i.e., nested within the abstract):
# author: Christopher L. Crawford,$^a$$^*$ He Yin,$^b$ Volker C. Radeloff,$^c$ and David S. Wilcove$^{a, d}$
# date: "`r format(Sys.time(), '%B %d, %Y')`"
# abstract: |
#   | $^a$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ
#   | $^b$Department of Geography, Kent State University, Kent, OH
#   | $^c$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI
#   | $^d$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ
#   |
#   | $^*$Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ




# formatted for LaTeX
# author:
#   - Christopher L. Crawford$^1$^[Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ]
#   - He Yin$^2$
#   - Alex Wiebe$^{4}$
#   - Volker C. Radeloff$^3$
#   - David S. Wilcove$^{1,4}$
# date: 
#   $^1$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ\newline
#   $^2$Department of Geography, Kent State University, Kent, OH\newline
#   $^3$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI\newline
#   $^4$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ\newline
#   \newline
#   \today
```

```{r load-packages-functions, eval = TRUE, echo = FALSE, include = FALSE}
source("/Users/christophercrawford/work/projects/biodiversity_abn/scripts/0_start.R")
```

```{r preview-chapter, eval=FALSE, include = FALSE}
rmarkdown::render("scripts/biodiv_abn_MS.Rmd",
                  output_file = 
                    paste0(p_proj, "writing/previews/", "Crawford_bd_abn_MS_draft",
                           format(Sys.Date(), ("_%Y_%m_%d")), ".docx"))

# Save final PDF with SI for submission:
# rmarkdown::render("scripts/Crawford_biodiv_abn_MS+SI.Rmd",
#                   output_file = paste0(p_proj, "writing/", "Crawford_biodiv_abn_MS+SI",
#                            format(Sys.Date(), ("_%Y_%m_%d")), ".pdf"),
#                   output_format = "bookdown::pdf_document2")

# bookdown::preview_chapter(
#   input = "abn_trajectories_ms.Rmd",
#   output_format = "bookdown::word_document2",
#   output_file = paste0("abn_duration_draft", format(Sys.Date(), ("_%Y_%m_%d")), ".docx"),
#   output_dir = paste0(p_proj, "writing/previews"),
#   output_options = 
#     list(reference_docx = "/Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx"),
#   preview = TRUE)
```

```{r set-chunk-opts, include = FALSE}
# set default chunk options
knitr::opts_chunk$set(
  out.width = "1\\textwidth",  # if compiling a pdf, include this option
  # comment = NA,
  # fig.width=6, fig.height=6, 
  echo = FALSE # the code itself will NOT be printed with results
  )

options(knitr.table.format = function() {
  if (knitr::is_latex_output())
    "latex" else "simple"
})
```

```{r produce-package-bib, include = FALSE, echo = FALSE}
# needed_packages
# github_packages
knitr::write_bib(x = c(needed_packages, github_packages), file = 'packages.bib')
# knitr::write_bib(x = c("car"), file = 'extra.bib')

# citation(package = needed_packages[1])

# knitr::write_bib(x = c(needed_packages[c(14)]), file = 'packages_top.bib')
# knitr::write_bib(x = c(needed_packages[3]), file = 'terra.bib')

# 
# needed_packages[1]
# packageDescription(needed_packages[1])
# 
# tinytex::tlmgr_install(
#   # pkgs = "ragged2e"
#   # pkgs = "standalone"
#   pkgs = "preview"
# )
```

```{r load-files, include = FALSE, eval = FALSE}

# set run:
run_label
aoh_run_date

# Additional paths --------------------------------------------------- #
p_plots <- "/Users/christophercrawford/work/projects/abandonment_trajectories/output/plots/"
p_derived2 <- paste0(p_dat_derived, run_label, "/derived_data/")

# load general spatial data on site locations --------------------------------------------------- #
site_df <- read.csv(file = paste0(p_dat_derived, "site_df.csv"))
site_df_w_size <- read.csv(file = paste0(p_dat_derived, "site_df_w_size.csv"))
site_sf <- st_read(paste0(p_dat_derived, "site_sf.shp"))


# load all length data (distilled)
length_distill_df <- read_csv(file = paste0(p_derived2, 
                                            "length_distill_df", run_label, ".csv"))

# load mean length data
mean_length_df <- read_csv(file = paste0(p_derived2, "mean_length_df", run_label, ".csv"))


# load summary statistic derived from lengths
summary_stats_all_sites <- read_csv(file = paste0(p_derived2, "summary_stats_all_sites", run_label, ".csv"))
summary_stats_all_sites_pooled <- read_csv(file = paste0(p_derived2, "summary_stats_all_sites_pooled", run_label, ".csv"))

# load area summary stats:
area_summary_df <- read_csv(file = paste0(p_derived2, "area_summary_df", run_label, ".csv"))
area_recult_threshold <- read_csv(file = paste0(p_derived2, "area_recult_threshold", run_label, ".csv"))


# load stats on proportion of abandoned cropland by land cover
abn_lc_area_2017 <- read_csv(file = paste0(p_derived2, "abn_lc_area_2017", run_label, ".csv"))
abn_prop_lc_2017 <- read_csv(file = paste0(p_derived2, "abn_prop_lc_2017", run_label, ".csv"))


# load combined persistence, area, and turnover datasets:
# also known as dat_l in "4_decay_models.Rmd"

persistence_dat <- 
  # dat_l <- 
  read_csv(file = paste0(p_derived2, "persistence_dat", run_label, ".csv")) %>%
  mutate(site = as.factor(site),
         bins = as.factor(bins),
         year_abn_bins = as.factor(year_abn_bins),
         cohort = as.factor(cohort))

area_dat <- read_csv(file = paste0(p_derived2, "area_dat", run_label, ".csv"))
turnover_dat <- read_csv(file = paste0(p_derived2, "turnover_dat", run_label, ".csv"))



# load decay rate data
load(file = paste0(p_derived2, "/decay_mod_coef_dfs.rds"), 
     verbose = TRUE)

# load extrapolation data
extrapolate_df_l3 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3", run_label, ".csv"))
extrapolate_df_l3_a2 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3_a2", run_label, ".csv"))


# load data for shaanxi maps --------------------------------------------------- #

# load data for decay models --------------------------------------------------- #
fitted_df_l3 <- read_csv(file = paste0(p_derived2, "/decay_mod_fitted_df_l3", run_label, ".csv"))

load(file = paste0(p_derived2, "/decay_mod_coef_dfs.rds"), verbose = TRUE)
# contains:
# coef_df_l3, coef_df_wide_l3, mean_coef_df_l3, mean_coef_df_wide_l3, mean_coef_proj_l3, time_to_l3


# load extrapolation data
extrapolate_df_l3 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3", run_label, ".csv"))

future_df_l3 <- read_csv(file = paste0(p_derived2, "/future_df_l3", run_label, ".csv"))

```

Preparing manuscript for [*Conservation Letters*](https://conbio.onlinelibrary.wiley.com/hub/journal/1755263x/homepage/forauthors.html):   
- 3000 word limit
- No more than 6 figures and tables
- 40 References

Structured as follows:  

1. Introduction. The Introduction should summarize briefly the background and aims, and end with a brief statement of what has been achieved by the work.
2. Methods. This section should contain sufficient detail so that all procedures can be clearly understood by the reader. Further technical details can be given in Supporting Information, which should be detailed enough that all procedures can be repeated (in conjunction with cited references).
3. Results. The Results section should present the evidence that supports the conclusions to be drawn in the Discussion. The Results section should conform to a high standard of rigour. Extended lines of inference, arguments or speculations should not be placed in the Results.
4. Discussion. The Discussion section should be separate from the Results section. It allows authors to propose their interpretation of the results, and to suggest what they mean in a wider context. It should end with a clear statement of the main conclusions of the research, and a clear explanation of their importance for conservation policy or practice.
5. Acknowledgements and Data. Acknowledgements of financial or institutional support should be brief (100 words maximum). In this section, authors are also expected to disclose any potential conflicts of interest, financial or otherwise, that a reasonable person could construe as possibly influencing the objectivity of the report. The absence of any statement is presumed to mean that there is nothing to disclose. Authors must provide a data accessibility statement, including a link to the repository they have used to archive their data (see policy on Data Sharing and Accessibility). They should also (where relevant) provide a statement of the ethics review process that was adhered to in gathering the data.
6. References


# Outline

Pieces

# Abstract (150 word limit) {.unnumbered}

Background/Question/Methods:

Although agricultural expansion and deforestation continues to be a dominant trend in many regions, tens of millions of hectares of croplands and pastures have simultaneously been abandoned as a result of shifting demographics, sociopolitical upheaval, environmental changes, and globalization. 
These abandoned agricultural lands increasingly represent opportunities for the restoration of ecosystems, with large potential environmental benefits in terms of carbon sequestration and increases in habitat availability for wildlife.
However, the extent to which abandonment presents an opportunity, or a threat, to biodiversity remains a topic of hot debate, due to varying species habitat requirements, threat status, and other factors.
Here, we combine a recent high-resolution annual land cover time series to estimate changes in area of habitat for bird and mammal species at 11 sites distributed across four continents.
We use annual land cover maps to identify cropland abandonment, and specifically track the age of abandoned fields.
We then calculate area of habitat (AOH) for bird and mammal species, by reconciling land cover classes to match IUCN habitat types, then excluding all land cover and elevations that are deemed unsuitable for a given species, according to the IUCN.

We 
We then explore how habitat area changed as a result of cropland abandonment by delving into differences between groups of species, looking at IUCN threat status, habitat preferences, habitat age requirements (early successional vs. mature habitat only)

We find that a significant portion of abandoned croplands are recultivated by the end of our time series, substantially curtailing the potential gains in AOH for different species.
We also find that 

Our land cover maps indicate that 
We show

We also investigate how recent abandonment and subsequent regeneration has affected landscape fragmentation. 

This time series shows that while 

agricultural abandonment , increasing levels of abandonment have raised hopes for recovery in some places.

Agricultural expansion continues to be the do

## Keywords {.unnumbered}

# Introduction {.unnumbered}

Three-four paragraphs to start

# Methods {.unnumbered}

Data

Analysis

## Calculating Area of Habitat

## Species range maps and filtering steps

We filtered species habitat associations based on season-specific habitat suitabilities provided in IUCN assessments for each individual species.

### Passage areas

Passage areas were excluded.

### Filtering species by habitat age requirements

Species were given a binary classification indicating whether it required mature habitat in at least one season at each site (in our case, any habitat requiring at least 30 years of secondary succession to become suitable for that species), which we used to further classify our results.
Classifications were manually determined based on species accounts from the handbook of the birds of the world, and handbook of the mammals of the world.
Neotropical bird species were independently verified by expert observer Alex Wiebe.


Note to self: marking species that I'm less sure about with 0.1 (on the cusp, but likely 0, non-mature habitat obligates) or 0.9 (on the cusp, but likely 1, mature habitat obligates).


## Mapping cropland abandonment and habitat

### Crosswalking land cover classess with IUCN habitat types

We also tested alternative ways to assign different IUCN habitat types to our coarser categories of forest and grassland.
Link to figure showing crosswalk Jaccard similarities. 



## Analysis

Linear models


# Results {.unnumbered}

Outline:  
1. Overall
2. By site
3. By IUCN threat status
4. AOH types
5. Observed vs. potential abandonment
6. Mature forest obligates. 

(ref:caption-abn-lc-2017) Land cover as of the year 2017, shown for all pixels that were abandoned at least once during the time series.

```{r abn-lc-2017, fig.cap = '(ref:caption-abn-lc-2017)'}
knitr::include_graphics(path = paste0(p_plots, "abn_land_cover_2017.pdf"))
```


(ref:caption-trend-overall-summary) TBD.

```{r aoh-trend-overall-summary, fig.cap = '(ref:caption-trend-overall-summary)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_by_vert_aoh_type", aoh_run_date,".pdf"))
```


(ref:caption-aoh-by-site-crop-abn) TBD.

```{r aoh-by-site-crop-abn, fig.cap = '(ref:caption-aoh-by-site-crop-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_site_crop_abn_iucn",
                                      aoh_run_date,".pdf"))
```


(ref:caption-aoh-by-iucn) TBD.

```{r aoh-by-iucn, fig.cap = '(ref:caption-aoh-by-iucn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_IUCN_status_lumped",
                                      aoh_run_date,".pdf"))
```


(ref:caption-aoh-trend-lines-crop-abn) Trend lines.

```{r aoh-trend-lines-crop-abn, fig.cap = '(ref:caption-aoh-trend-lines-crop-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_crop_abn_iucn",
                                      aoh_run_date,".pdf"))
```


(ref:caption-aoh-obs-v-pot) Proportion of species that gain or lose AOH as a result of abandonment during our time series, for both observed abandonment and a scenario of potential abandonment, assuming no recultivation.

```{r aoh-obs-v-pot, fig.cap = '(ref:caption-aoh-obs-v-pot)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_obs_v_pot_overall",
                                      aoh_run_date,".pdf"))
```




# Discussion {.unnumbered}

# Acknowledgements {.unnumbered}

## Conflict of Interest {.unnumbered}

The authors declare no conflicts of interest.

## Ethics and Funding {.unnumbered}

## Data Availability Statement {.unnumbered}

All datasets generated and analyzed for this article are archived at Zenodo (URL TBD). 
The annual land cover maps that underlie our analysis are also archived at Zenodo (URL TBD, doi:), along with derived maps of abandonment age and derived datasets summarizing area, persistence, and turnover of abandonment.
Code to replicate these analyses is available on GitHub at <https://github.com/chriscra/biodiversity_abandonment>, and is archived at Zenodo (URL TBD).

# References {.unnumbered}

BirdLife International and Handbook of the Birds of the World (2021) Bird species distribution maps of the world. Version 2021.1. Available at http://datazone.birdlife.org/species/requestdis.

Request blurb:
This IUCN Assessment data will be used for a dissertation research project designed to investigate the impact of agricultural abandonment on biodiversity. The final product will be academic journal articles, which will be shared with the IUCN science team upon completion. Please don't hesitate to get in touch if you have any questions (ccrawford@princeton.edu). Thank you! - Christopher Crawford, PhD Candidate, Princeton University.

Range maps downloaded on January 4th, 2022 (amp, mam, reptiles, from IUCN, and birds from BirdLife International)

::: {#refs}
:::

\newpage

\beginsupplement

# Supplementary information {.unnumbered}

\tableofcontents
\listoftables
\listoffigures

\newpage

# Extended Results {#results-si}

# Extended Methods

# Supplemental Figures

(ref:caption-richness-by-site) Species richness at each site, or the count of species with some Area of Habitat (AOH) at each site. Richness is shown for threatened ("Vulnerable," "Endangered," or "Critically Endangered") or non-threatened ("Least Concern", "Near Threatened"), as assessed by the IUCN. Species classified as "Data Deficient" are excluded.

```{r richness-by-site, fig.cap = '(ref:caption-richness-by-site)'}
knitr::include_graphics(path = paste0(
  p_plots, "richness_by_site_w_category_facet_color.pdf"))
```



# Supplemental Tables

