---
title: "Cropland abandonment benefits a majority of birds and mammals, but rarely compensates for habitat loss"
author: Christopher L. Crawford,$^a$$^*$, R. Alex Wiebe,$^{b}$ He Yin,$^c$ Volker C. Radeloff,$^d$ and David S. Wilcove$^{a, b}$
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: |
  | $^a$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ
  | $^b$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ
  | $^c$Department of Geography, Kent State University, Kent, OH
  | $^d$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI
  |
  | $^*$Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ
output:
  bookdown::html_document2: 
    self_contained: true
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: false
    keep_tex: false
    number_sections: true
    includes:
      in_header: "preamble.tex"
  bookdown::word_document2:
    reference_docx: /Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx
    number_sections: false
mainfont: Times New Roman
fontsize: 11pt
# linestretch: 1.25
# documentclass: report
editor_options:
  chunk_output_type: console
bibliography: [/Users/christophercrawford/work/library/library.bib, packages.bib]
link-citations: yes
csl: conservation-letters.csl
urlcolor: blue
linkcolor: blue
citecolor: blue
---
```{r title-possibilities, include = FALSE}
# Title possibilities:
# 
# - Cropland abandonment benefits more birds and mammals than it harms, but rarely compensates for habitat loss
# - Cropland abandonment benefits more birds and mammals than it harms, but requires protection in order to compensate for habitat loss
```

```{r misc-notes, include = FALSE, eval = FALSE}
# Authors and Affiliations, formatted to look nice in Word (i.e., nested within the abstract):
# author: Christopher L. Crawford,$^a$$^*$ He Yin,$^b$ Volker C. Radeloff,$^c$ and David S. Wilcove$^{a, d}$
# date: "`r format(Sys.time(), '%B %d, %Y')`"
# abstract: |
#   | $^a$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ
#   | $^b$Department of Geography, Kent State University, Kent, OH
#   | $^c$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI
#   | $^d$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ
#   |
#   | $^*$Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ




# formatted for LaTeX
# author:
#   - Christopher L. Crawford$^1$^[Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ]
#   - He Yin$^2$
#   - Alex Wiebe$^{4}$
#   - Volker C. Radeloff$^3$
#   - David S. Wilcove$^{1,4}$
# date: 
#   $^1$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ\newline
#   $^2$Department of Geography, Kent State University, Kent, OH\newline
#   $^3$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI\newline
#   $^4$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ\newline
#   \newline
#   \today
```

```{r initialize, eval = TRUE, echo = FALSE, include = FALSE}
source("/Users/christophercrawford/work/projects/biodiversity_abn/scripts/0_start.R")
source("/Users/christophercrawford/work/projects/biodiversity_abn/scripts/_util/_util_files.R")
# my_format <- "simple"

```

```{r preview-chapter, eval=FALSE, include = FALSE}
my_format <- "simple"
rmarkdown::render("scripts/biodiv_abn_MS.Rmd",
                  output_file = 
                    paste0(p_proj, "writing/previews/", "Crawford_bd_abn_MS_draft",
                           format(Sys.Date(), ("_%Y_%m_%d")), ".docx"),
                  output_format = bookdown::word_document2(
                    reference_docx = "/Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx",
                    number_sections = FALSE))

# Save final PDF with SI for submission:
my_format <- "latex"
rmarkdown::render("scripts/biodiv_abn_MS.Rmd",
                  output_file = paste0(p_proj, "writing/previews/", "Crawford_bd_abn_MS+SI",
                           format(Sys.Date(), ("_%Y_%m_%d")), ".pdf"),
                  output_format = "bookdown::pdf_document2")


# html
# my_format <- "simple"
# rmarkdown::render("scripts/biodiv_abn_MS.Rmd",
#                   output_file = 
#                     paste0(p_proj, "writing/previews/", "Crawford_bd_abn_MS_draft",
#                            format(Sys.Date(), ("_%Y_%m_%d")), ".html"),
#                   # output_format = "bookdown::html_document2",
#                   html_document2(self_contained = TRUE))


# bookdown::preview_chapter(
#   input = "abn_trajectories_ms.Rmd",
#   output_format = "bookdown::word_document2",
#   output_file = paste0("abn_duration_draft", format(Sys.Date(), ("_%Y_%m_%d")), ".docx"),
#   output_dir = paste0(p_proj, "writing/previews"),
#   output_options = 
#     list(reference_docx = "/Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx"),
#   preview = TRUE)
```

```{r set-chunk-opts, include = FALSE}
# set default chunk options
knitr::opts_chunk$set(
  # out.width = "1\\textwidth",  # if compiling a pdf, include this option
  # comment = NA,
  # fig.width=6, fig.height=6, 
  warning = FALSE, # hide warnings from code chunks
  echo = FALSE # the code itself will NOT be printed with results
  )

options(knitr.table.format = function() {
  if (knitr::is_latex_output())
    "latex" else "simple"
})
```

```{r produce-package-bib, eval = FALSE, include = FALSE, echo = FALSE}
# needed_packages
# github_packages
knitr::write_bib(x = c(needed_packages, github_packages), file = 'packages.bib')
# knitr::write_bib(x = c("car"), file = 'extra.bib')

# citation(package = needed_packages[1])

# knitr::write_bib(x = c(needed_packages[c(14)]), file = 'packages_top.bib')
# knitr::write_bib(x = c(needed_packages[3]), file = 'terra.bib')

# 
# needed_packages[1]
# packageDescription(needed_packages[1])
# 
# tinytex::tlmgr_install(
#   # pkgs = "ragged2e"
#   # pkgs = "standalone"
#   pkgs = "preview"
# )
```

```{r abstract-stats, include = FALSE}
abs_aoh_tmp <- 
  aoh_trends_by_sp %>%
  filter(passage_type == "exclude_passage",
         vert_class != "amp", 
         !aoh_type %in% c("abn_iucn", "potential_abn_iucn"),
         # !grepl("potential", aoh_type),
         !(vert_class == "bird" & mature_forest_obl > 0.5)
         ) %>%
  group_by(aoh_type, vert_class, passage_type,
           overall_trend,
           trend_direction) %>%
  summarise(n_sp = n()) %>%
  arrange(vert_class, aoh_type) %>%
  # filter(aoh_type == "crop_abn_iucn") %>%
  ungroup() %>%
  group_by(aoh_type, vert_class) %>%
  mutate(total_sp = sum(n_sp),
         p = 100 * n_sp/total_sp,
         pr = round(p, digits = 2),
         pp = prettyNum(pr, nsmall = 2))


abstract_stats <- list(
  percent_winner_bird_obs = abs_aoh_tmp %>% 
    filter(aoh_type == "crop_abn_iucn", vert_class == "bird") %>%
    slice_max(p) %>% .$pp,
  
  percent_winner_bird_pot = abs_aoh_tmp %>% 
    filter(aoh_type == "crop_abn_potential_iucn", vert_class == "bird") %>%
    slice_max(p) %>% .$pp,
  
  percent_winner_mam_obs = abs_aoh_tmp %>% 
    filter(aoh_type == "crop_abn_iucn", vert_class == "mam") %>%
    slice_max(p) %>% .$pp,
  
  percent_winner_mam_pot = abs_aoh_tmp %>% 
    filter(aoh_type == "crop_abn_potential_iucn", vert_class == "mam") %>%
    slice_max(p) %>% .$pp,
  
  percent_winner_bird_max_obs = abs_aoh_tmp %>% 
    filter(aoh_type == "max_abn_iucn", vert_class == "bird") %>%
    slice_max(p) %>% .$pp,
  
  percent_winner_mam_max_obs = abs_aoh_tmp %>% 
    filter(aoh_type == "max_abn_iucn", vert_class == "mam") %>%
    slice_max(p) %>% .$pp,
  
  percent_winner_bird_full_obs = abs_aoh_tmp %>% 
    filter(aoh_type == "full_iucn", vert_class == "bird") %>%
    slice_max(p) %>% .$pp,
  
  percent_winner_mam_full_obs = abs_aoh_tmp %>% 
    filter(aoh_type == "full_iucn", vert_class == "mam") %>%
    slice_max(p) %>% .$pp

)


aoh_trends_by_sp %>%
  filter(passage_type == "exclude_passage",
         vert_class != "amp", 
         !aoh_type %in% c("abn_iucn", "potential_abn_iucn"),
         # !grepl("potential", aoh_type),
         !(vert_class == "bird" & mature_forest_obl > 0.5)
         ) %>%
  group_by(aoh_type, vert_class, passage_type,
           # overall_trend, 
           trend_direction
           ) %>%
  summarise(n_sp = n()) %>%
  arrange(vert_class, aoh_type) %>%
  # filter(aoh_type == "crop_abn_iucn") %>%
  ungroup() %>%
  group_by(aoh_type, vert_class) %>%
  mutate(total_sp = sum(n_sp),
         p = 100 * n_sp/total_sp,
         pr = round(p, digits = 2),
         pp = prettyNum(pr, nsmall = 2)) %>%
  filter(trend_direction == "gain") %>%
  select(aoh_type, vert_class, n_sp, total_sp, pr)
  # slice_max(p) %>% .$pp
  # print(n = 50)

```

```{r get-stats-aoh-stats-main, include = FALSE}
# number of small-ranged birds and mammals
aoh_trends_by_sp %>% 
  filter(passage_type == "exclude_passage",
         vert_class != "amp",
         !(vert_class == "bird" & mature_forest_obl > 0.5),
         aoh_type == "crop_abn_iucn",
         range_size_quantile < 0.5)

aoh_stat_tmp_by_site_crop <-
  aoh_trends %>%
      filter(!is.na(trend),
             passage_type == "exclude_passage",
             vert_class != "amp",
             !(vert_class == "bird" & mature_forest_obl > 0.5),
             aoh_type == "crop_abn_iucn"
             # redlistCategory != "Data Deficient"
             ) %>%
      group_by(aoh_type, vert_class, passage_type, 
               site,
               trend) %>%
      summarise(n_sp = n()) %>% ungroup() %>%
      arrange(vert_class, aoh_type) %>%
  pivot_wider(names_from = trend, values_from = n_sp) %>%
  mutate(prop_wl = gain/loss) %>% 
  select(vert_class, site, prop_wl) %>% 
  arrange(vert_class, prop_wl) 

aoh_stat_tmp_by_site_crop %>% print(n = 22) 
aoh_stat_tmp_by_site_crop %>% group_by(vert_class) %>% slice_min(prop_wl)

# mean by site
aoh_stat_tmp_by_site_crop %>% group_by(vert_class) %>% summarise(mean_prop_wl = mean(prop_wl))

aoh_overall_tmp <- 
  aoh_trends_by_sp %>%
  filter(passage_type == "exclude_passage",
         vert_class != "amp", 
         !aoh_type %in% c("abn_iucn", "potential_abn_iucn"),
         !grepl("potential", aoh_type),
         !(vert_class == "bird" & mature_forest_obl > 0.5)
         ) %>%
  group_by(aoh_type, vert_class, passage_type,
           overall_trend, trend_direction, trend_consistency) %>%
  summarise(n_sp = n()) %>%
  arrange(vert_class, aoh_type)

aoh_stats_main <- list(
  n_gain_bird = filter(aoh_overall_tmp, 
                       aoh_type == "crop_abn_iucn", overall_trend == "gain",
                       vert_class == "bird")$n_sp,
  n_gain_mam = filter(aoh_overall_tmp, 
                       aoh_type == "crop_abn_iucn", overall_trend == "gain",
                       vert_class == "mam")$n_sp,
  n_loss_bird = filter(aoh_overall_tmp, 
                       aoh_type == "crop_abn_iucn", overall_trend == "loss",
                       vert_class == "bird")$n_sp,
  n_loss_mam = filter(aoh_overall_tmp, 
                       aoh_type == "crop_abn_iucn", overall_trend == "loss",
                       vert_class == "mam")$n_sp,
  wl_prop_bird_site_sp = aoh_trends %>%
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "crop_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    arrange(vert_class, aoh_type) %>%
    pivot_wider(names_from = trend, values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_site_sp = aoh_trends %>%
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "crop_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    arrange(vert_class, aoh_type) %>%
    pivot_wider(names_from = trend, values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_bird_sp = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "crop_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             overall_trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_bird_sp_w_weak = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "crop_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             trend_direction) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_sp = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "crop_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             overall_trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_sp_w_weak = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "crop_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             trend_direction) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_site_min = aoh_stat_tmp_by_site_crop %>% filter(site !="iraq") %>% 
    # group_by(vert_class) %>% 
    slice_min(prop_wl) %>% .$prop_wl %>% round(digits = 2),
  
  wl_prop_site_max = aoh_stat_tmp_by_site_crop %>% filter(site !="iraq") %>% 
    # group_by(vert_class) %>% 
    slice_max(prop_wl) %>% .$prop_wl %>% round(digits = 2),
  
  wl_iraq = round(aoh_stat_tmp_by_site_crop %>% filter(site == "iraq", vert_class == "bird") %>%
    .$prop_wl, digits = 2)
)

```
```{r get-stats-obs-v-pot, include = FALSE}
# Two files
aoh_p_change_obs_v_pot # and
aoh_p_change_obs_v_pot_summary

# set up tmp files:
mg_improve_tmp <- aoh_p_change_obs_v_pot_summary %>% filter(site == "mato_grosso", change_direction == "Improve")

# set up stats:
obs_v_pot_stats <- list(
  mg_n_winner_bird_obs = aoh_trends %>%
  filter(passage_type == "exclude_passage", vert_class != "amp", 
         grepl("crop", aoh_type),
         !(vert_class == "bird" & mature_forest_obl > 0.5)) %>%
  group_by(aoh_type, vert_class, passage_type, 
           site, trend) %>%
  summarise(n_sp = n()) %>%
  arrange(vert_class, trend, aoh_type) %>%
  filter(site == "mato_grosso", trend == "gain") %>%
  pivot_wider(values_from = n_sp, names_from = aoh_type) %>%
  filter(vert_class == "bird") %>% .$crop_abn_iucn,
  
  mg_n_winner_bird_pot = aoh_trends %>%
  filter(passage_type == "exclude_passage",
         vert_class != "amp", grepl("crop", aoh_type),
         !(vert_class == "bird" & mature_forest_obl > 0.5)) %>%
  group_by(aoh_type, vert_class, passage_type, 
           site, trend) %>%
  summarise(n_sp = n()) %>%
  arrange(vert_class, trend, aoh_type) %>%
  filter(site == "mato_grosso", trend == "gain") %>%
  pivot_wider(values_from = n_sp, names_from = aoh_type) %>%
  filter(vert_class == "bird") %>% .$crop_abn_potential_iucn,
  
  mg_n_winner_mam_obs = aoh_trends %>%
  filter(passage_type == "exclude_passage", vert_class != "amp", 
         grepl("crop", aoh_type),
         !(vert_class == "bird" & mature_forest_obl > 0.5)) %>%
  group_by(aoh_type, vert_class, passage_type, 
           site, trend) %>%
  summarise(n_sp = n()) %>%
  arrange(vert_class, trend, aoh_type) %>%
  filter(site == "mato_grosso", trend == "gain") %>%
  pivot_wider(values_from = n_sp, names_from = aoh_type) %>%
  filter(vert_class == "mam") %>% .$crop_abn_iucn,
  
  mg_n_winner_mam_pot = aoh_trends %>%
  filter(passage_type == "exclude_passage",
         vert_class != "amp", grepl("crop", aoh_type),
         !(vert_class == "bird" & mature_forest_obl > 0.5)) %>%
  group_by(aoh_type, vert_class, passage_type, 
           site, trend) %>%
  summarise(n_sp = n()) %>%
  arrange(vert_class, trend, aoh_type) %>%
  filter(site == "mato_grosso", trend == "gain") %>%
  pivot_wider(values_from = n_sp, names_from = aoh_type) %>%
  filter(vert_class == "mam") %>% .$crop_abn_potential_iucn,
  
  
  mg_improve_n_total = mg_improve_tmp %>% summarise(n = sum(n_sp_in_group)),
  mg_improve_n_bird_no_trend = mg_improve_tmp %>% 
    filter(str_detect(trend_change, "no"), vert_class == "bird") %>% .$n_sp_in_group,
  
  mg_improve_n_mam_no_trend = mg_improve_tmp %>% 
    filter(str_detect(trend_change, "no"), vert_class == "mam") %>% .$n_sp_in_group,
  
  mg_improve_n_bird_loss = mg_improve_tmp %>% 
    filter(str_detect(trend_change, "loss"), vert_class == "bird") %>% .$n_sp_in_group,
  
  mg_improve_n_mam_loss = mg_improve_tmp %>% 
    filter(str_detect(trend_change, "loss"), vert_class == "mam") %>% .$n_sp_in_group,
  
  mg_improve_mean_p_gain = 
    format(round(
       mg_improve_tmp %>%
            group_by(change_direction) %>% 
            summarise(n_sp = sum(n_sp_in_group),
                      w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>% 
            .$w_mean, digits = 2), nsmall = 2),
  
  mg_worsen_n_bird = aoh_p_change_obs_v_pot_summary %>% 
        filter(site == "mato_grosso", change_direction == "Worsen",
               vert_class == "bird") %>% .$n_sp_in_group,
  mg_worsen_n_mam = aoh_p_change_obs_v_pot_summary %>% 
        filter(site == "mato_grosso", change_direction == "Worsen",
               vert_class == "mam") %>% .$n_sp_in_group,
  
  mg_worsen_mean_p_loss = format(round(aoh_p_change_obs_v_pot_summary %>% 
        filter(site == "mato_grosso", change_direction == "Worsen") %>%
            group_by(change_direction) %>% 
            summarise(n_sp = sum(n_sp_in_group),
                      w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>% 
            .$w_mean * -1, digits = 2), nsmall = 2),
  
  # how many species improved trend?
  n_sp_improve_non_mg_bird = aoh_p_change_obs_v_pot %>%
    select(vert_class, site, binomial, contains("change")) %>%
    filter(change_direction == "Improve", site != "mato_grosso") %>%
    select(vert_class, binomial, trend_change, change_direction) %>% unique() %>% 
    group_by(vert_class, change_direction) %>%
    summarise(n_sp = length(unique(binomial))) %>%
    filter(vert_class == "bird") %>% .$n_sp,
  
  n_sp_improve_non_mg_mam = aoh_p_change_obs_v_pot %>%
    select(vert_class, site, binomial, contains("change")) %>%
    filter(change_direction == "Improve", site != "mato_grosso") %>%
    select(vert_class, binomial, trend_change, change_direction) %>% unique() %>% 
    group_by(vert_class, change_direction) %>%
    summarise(n_sp = length(unique(binomial))) %>%
    filter(vert_class == "mam") %>% .$n_sp,
  
  n_same_loss_bird = 
    aoh_p_change_obs_v_pot_summary %>%
    group_by(vert_class, trend_change, change_direction) %>% 
    summarise(n_sp = sum(n_sp_in_group),
              w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>%
    arrange(vert_class, change_direction) %>%
    filter(trend_change == "same (loss)", vert_class == "bird") %>% .$n_sp,
  
  n_same_loss_mam = 
    aoh_p_change_obs_v_pot_summary %>%
    group_by(vert_class, trend_change, change_direction) %>% 
    summarise(n_sp = sum(n_sp_in_group),
              w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>%
    arrange(vert_class, change_direction) %>%
    filter(trend_change == "same (loss)", vert_class == "mam") %>% .$n_sp,
  
  mean_p_loss_same = round(
    aoh_p_change_obs_v_pot_summary %>%
    group_by(vert_class, trend_change, change_direction) %>% 
    summarise(n_sp = sum(n_sp_in_group),
              w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>%
    arrange(vert_class, change_direction) %>% ungroup() %>%
    filter(trend_change == "same (loss)") %>% 
    summarise(wm = sum(n_sp*w_mean) / sum(n_sp)) %>% 
    .$wm * -1, digits = 2),
  
  mean_p_loss_site_min = round(
    aoh_p_change_obs_v_pot_summary %>%
    filter(trend_change == "same (loss)") %>% 
    ungroup() %>% slice_max(mean_p_change) %>%
      .$mean_p_change * -1, digits = 2),
  
  mean_p_loss_site_max = round(aoh_p_change_obs_v_pot_summary %>%
    filter(trend_change == "same (loss)") %>% 
    ungroup() %>% slice_min(mean_p_change) %>% 
      .$mean_p_change * -1, digits = 2),
  
  mean_p_loss_site_min_bird = round(
    aoh_p_change_obs_v_pot_summary %>%
      filter(trend_change == "same (loss)", vert_class == "bird") %>%
      group_by(vert_class) %>% slice_max(mean_p_change) %>% 
      .$mean_p_change * -1, digits = 2),
  
  mean_p_loss_site_max_bird = round(
    aoh_p_change_obs_v_pot_summary %>% 
      filter(trend_change == "same (loss)", vert_class == "bird") %>%
      group_by(vert_class) %>% slice_min(mean_p_change) %>% 
      .$mean_p_change * -1, digits = 2),
  
  mean_p_loss_site_min_mam = round(
    aoh_p_change_obs_v_pot_summary %>%
      filter(trend_change == "same (loss)", vert_class == "mam") %>%
      group_by(vert_class) %>% slice_max(mean_p_change) %>%
      .$mean_p_change * -1, digits = 2),
  
  mean_p_loss_site_max_mam = round(
    aoh_p_change_obs_v_pot_summary %>%
      filter(trend_change == "same (loss)", vert_class == "mam") %>%
      group_by(vert_class) %>% slice_min(mean_p_change) %>% 
      .$mean_p_change * -1, digits = 2),
  
  
  n_same_gain_bird = 
    aoh_p_change_obs_v_pot_summary %>%
    group_by(vert_class, trend_change, change_direction) %>% 
    summarise(n_sp = sum(n_sp_in_group),
              w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>%
    arrange(vert_class, change_direction) %>%
    filter(trend_change == "same (gain)", vert_class == "bird") %>% .$n_sp,
  
  n_same_gain_mam = 
    aoh_p_change_obs_v_pot_summary %>%
    group_by(vert_class, trend_change, change_direction) %>% 
    summarise(n_sp = sum(n_sp_in_group),
              w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>%
    arrange(vert_class, change_direction) %>%
    filter(trend_change == "same (gain)", vert_class == "mam") %>% .$n_sp,
  
  mean_p_gain_same = aoh_p_change_obs_v_pot_summary %>%
    group_by(vert_class, trend_change, change_direction) %>% 
    summarise(n_sp = sum(n_sp_in_group),
              w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>%
    arrange(vert_class, change_direction) %>% ungroup() %>%
    filter(trend_change == "same (gain)") %>% 
    summarise(wm = sum(n_sp*w_mean) / sum(n_sp)) %>% 
    .$wm %>% round(digits = 2),
  
  mean_p_gain_site_min = aoh_p_change_obs_v_pot_summary %>%
    filter(trend_change == "same (gain)") %>% 
    ungroup() %>% slice_min(mean_p_change) %>% .$mean_p_change %>%
    round(digits = 2),
  mean_p_gain_site_max = aoh_p_change_obs_v_pot_summary %>%
    filter(trend_change == "same (gain)") %>% 
    ungroup() %>% slice_max(mean_p_change) %>% .$mean_p_change %>%
    round(digits = 2)
  )


# ------------------------------------------------- #

# how many species improved trend?
aoh_p_change_obs_v_pot %>%
  select(vert_class, site, binomial, contains("change")) %>%
  filter(change_direction == "Improve", site == "mato_grosso") %>%
  select(vert_class, binomial, trend_change, change_direction) %>% unique() %>% 
  print(n = 30) %>%
  group_by(vert_class, change_direction) %>%
  summarise(n_sp = length(unique(binomial)))

# Number of species with various trends in each scenario:
aoh_trends %>%
  filter(passage_type == "exclude_passage",
         vert_class == "bird", grepl("crop", aoh_type),
         mature_forest_obl < 0.5) %>%
  group_by(aoh_type, vert_class, passage_type, site, trend) %>%
  summarise(n_sp = n(), ) %>% 
  filter(site != "mato_grosso") %>%
  arrange(trend)

# percent changes in area between scenarios
aoh_p_change_obs_v_pot_summary %>% print(n = 100)

aoh_p_change_obs_v_pot_summary %>% 
  filter(change_direction == "Improve", site != "mato_grosso") %>%
  group_by(vert_class) %>% summarise(sum(n_sp_in_group))

aoh_p_change_obs_v_pot_summary %>%
  group_by(vert_class, 
           trend_change, change_direction) %>% 
  summarise(n_sp = sum(n_sp_in_group),
            w_mean = sum(mean_p_change*n_sp_in_group) / n_sp) %>%
  arrange(vert_class, 
          change_direction)


aoh_p_change_obs_v_pot_summary %>% 
  filter(site == "mato_grosso")
```
```{r get-stats-aoh-overall, include = FALSE}
aoh_stat_tmp_by_site_max <-
  aoh_trends %>%
      filter(!is.na(trend),
             passage_type == "exclude_passage",
             vert_class != "amp",
             !(vert_class == "bird" & mature_forest_obl > 0.5),
             aoh_type == "max_abn_iucn"
             ) %>%
      group_by(aoh_type, vert_class, passage_type, 
               site,
               trend) %>%
      summarise(n_sp = n()) %>% ungroup() %>%
      arrange(vert_class, aoh_type) %>%
  pivot_wider(names_from = trend, values_from = n_sp) %>%
  mutate(prop_wl = gain/loss) %>% 
  select(vert_class, site, prop_wl) %>% 
  arrange(vert_class, prop_wl) 

aoh_stat_tmp_by_site_max %>% print(n = 22) 
aoh_stat_tmp_by_site_max %>% group_by(vert_class) %>% slice_min(prop_wl)

aoh_stats_max <- list(
  n_gain_bird = filter(aoh_overall_tmp, 
                       aoh_type == "max_abn_iucn", overall_trend == "gain",
                       vert_class == "bird")$n_sp,
  n_gain_mam = filter(aoh_overall_tmp, 
                       aoh_type == "max_abn_iucn", overall_trend == "gain",
                       vert_class == "mam")$n_sp,
  n_loss_bird = filter(aoh_overall_tmp, 
                       aoh_type == "max_abn_iucn", overall_trend == "loss",
                       vert_class == "bird")$n_sp,
  n_loss_mam = filter(aoh_overall_tmp, 
                       aoh_type == "max_abn_iucn", overall_trend == "loss",
                       vert_class == "mam")$n_sp,
  
  wl_prop_bird_site_sp = aoh_trends %>%
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "max_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    arrange(vert_class, aoh_type) %>%
    pivot_wider(names_from = trend, values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_site_sp = aoh_trends %>%
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "max_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    arrange(vert_class, aoh_type) %>%
    pivot_wider(names_from = trend, values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_bird_sp = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "max_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             overall_trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_bird_sp_w_weak = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "max_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             trend_direction) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_sp = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "max_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             overall_trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_sp_w_weak = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "max_abn_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             trend_direction) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_site_min = aoh_stat_tmp_by_site_max %>% filter(site !="iraq") %>% 
    # group_by(vert_class) %>% 
    slice_min(prop_wl) %>% .$prop_wl %>% round(digits = 2),
  
  wl_prop_site_max = aoh_stat_tmp_by_site_max %>% filter(site !="iraq") %>% 
    # group_by(vert_class) %>% 
    slice_max(prop_wl) %>% .$prop_wl %>% round(digits = 2),
  
  wl_iraq = round(aoh_stat_tmp_by_site_max %>% filter(site == "iraq", vert_class == "bird") %>%
    .$prop_wl, digits = 2)

)

# ------ entire landscape ------ #
aoh_stat_tmp_by_site_full <-
  aoh_trends %>%
      filter(!is.na(trend),
             passage_type == "exclude_passage",
             vert_class != "amp",
             !(vert_class == "bird" & mature_forest_obl > 0.5),
             aoh_type == "full_iucn"
             ) %>%
      group_by(aoh_type, vert_class, passage_type, 
               site,
               trend) %>%
      summarise(n_sp = n()) %>% ungroup() %>%
      arrange(vert_class, aoh_type) %>%
  pivot_wider(names_from = trend, values_from = n_sp) %>%
  mutate(prop_wl = gain/loss) %>% 
  select(vert_class, site, prop_wl) %>% 
  arrange(vert_class, prop_wl) 

aoh_stat_tmp_by_site_full %>% print(n = 22) 
aoh_stat_tmp_by_site_full %>% group_by(vert_class) %>% slice_min(prop_wl)

aoh_stats_full <- list(
  n_gain_bird = filter(aoh_overall_tmp, 
                       aoh_type == "full_iucn", overall_trend == "gain",
                       vert_class == "bird")$n_sp,
  n_gain_mam = filter(aoh_overall_tmp, 
                       aoh_type == "full_iucn", overall_trend == "gain",
                       vert_class == "mam")$n_sp,
  n_loss_bird = filter(aoh_overall_tmp, 
                       aoh_type == "full_iucn", overall_trend == "loss",
                       vert_class == "bird")$n_sp,
  n_loss_mam = filter(aoh_overall_tmp, 
                       aoh_type == "full_iucn", overall_trend == "loss",
                       vert_class == "mam")$n_sp,
  wl_prop_bird_site_sp = aoh_trends %>%
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "full_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    arrange(vert_class, aoh_type) %>%
    pivot_wider(names_from = trend, values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_site_sp = aoh_trends %>%
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "full_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    arrange(vert_class, aoh_type) %>%
    pivot_wider(names_from = trend, values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_bird_sp = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "full_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             overall_trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_bird_sp_w_weak = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "full_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             trend_direction) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "bird") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_sp = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "full_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             overall_trend) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_mam_sp_w_weak = aoh_trends_by_sp %>% 
    filter(passage_type == "exclude_passage",
           vert_class != "amp",
           !(vert_class == "bird" & mature_forest_obl > 0.5),
           aoh_type == "full_iucn") %>%
    group_by(aoh_type, vert_class, passage_type, 
             trend_direction) %>%
    summarise(n_sp = n()) %>% ungroup() %>%
    pivot_wider(names_from = contains("trend"), values_from = n_sp) %>%
    mutate(prop_wl = gain/loss) %>% filter(vert_class == "mam") %>%
    .$prop_wl %>% round(digits = 2),
  
  wl_prop_site_min = aoh_stat_tmp_by_site_full %>% filter(site !="iraq") %>% 
    # group_by(vert_class) %>% 
    slice_min(prop_wl) %>% .$prop_wl %>% round(digits = 2),
  
  wl_prop_site_max = aoh_stat_tmp_by_site_full %>% filter(site !="iraq") %>% 
    # group_by(vert_class) %>% 
    slice_max(prop_wl) %>% .$prop_wl %>% round(digits = 2),
  
  wl_iraq = round(aoh_stat_tmp_by_site_full %>% filter(site == "iraq", vert_class == "bird") %>%
    .$prop_wl, digits = 2)
)
```

# Abstract (150 word limit)^[Preparing manuscript for [*Conservation Letters*](https://conbio.onlinelibrary.wiley.com/hub/journal/1755263x/homepage/forauthors.html), with 3000 word limit, no more than 6 figures and tables, and 40 References.] {.unnumbered}

Although agricultural expansion continues to drive habitat loss in many regions, substantial areas of cropland have been abandoned due to demographic shifts, socioeconomic and technological changes, and globalization.
Abandoned lands increasingly represent opportunities to restore ecosystems yet spatial variation in the distribution of species and their habitats, along with a lack of detailed maps of abandonment, have resulted in uncertainty and controversy over whether abandonment will benefit, or harm, biodiversity.
Here, we use annual land cover maps to measure habitat area for 2023 bird and mammal species from 1987-2017 at eleven sites across four continents.
We find that the majority of bird (`r abstract_stats$percent_winner_bird_obs`%) and mammal (`r abstract_stats$percent_winner_mam_obs`%) species experience statistically significant gains in area of habitat as a direct result of cropland abandonment.
However, a greater share of species would have gained overall (`r abstract_stats$percent_winner_bird_pot`% and `r abstract_stats$percent_winner_mam_pot`% respectively) if no abandoned croplands had been recultivated.
Despite these encouraging gains, the majority of bird species had negative trends overall when accounting for a) habitat loss prior to cultivation or b) changes across the entire extent of each site (mammals had about equal winners and losers in both scenarios).
Thus, while abandonment presents an opportunity for most species, limiting recultivation and reducing habitat loss irrespective of abandonment remains critical for the future of biodiversity.

<!-- Structured as follows:   -->

<!-- 1. Introduction. The Introduction should summarize briefly the background and aims, and end with a brief statement of what has been achieved by the work. -->
<!-- 2. Methods. This section should contain sufficient detail so that all procedures can be clearly understood by the reader. Further technical details can be given in Supporting Information, which should be detailed enough that all procedures can be repeated (in conjunction with cited references). -->
<!-- 3. Results. The Results section should present the evidence that supports the conclusions to be drawn in the Discussion. The Results section should conform to a high standard of rigour. Extended lines of inference, arguments or speculations should not be placed in the Results. -->
<!-- 4. Discussion. The Discussion section should be separate from the Results section. It allows authors to propose their interpretation of the results, and to suggest what they mean in a wider context. It should end with a clear statement of the main conclusions of the research, and a clear explanation of their importance for conservation policy or practice. -->
<!-- 5. Acknowledgements and Data. Acknowledgements of financial or institutional support should be brief (100 words maximum). In this section, authors are also expected to disclose any potential conflicts of interest, financial or otherwise, that a reasonable person could construe as possibly influencing the objectivity of the report. The absence of any statement is presumed to mean that there is nothing to disclose. Authors must provide a data accessibility statement, including a link to the repository they have used to archive their data (see policy on Data Sharing and Accessibility). They should also (where relevant) provide a statement of the ethics review process that was adhered to in gathering the data. -->
<!-- 6. References -->

# Introduction {.unnumbered}

Amidst continued cropland expansion, hundreds of millions of hectares of agricultural lands around the world have been abandoned since 1950 due to socioeconomic, demographic, and technological changes [@ReyBenayas2007; @Li2017; @Potapov2022; @Chazdon2020].
While agricultural expansion and associated habitat loss remains the foremost threat to biodiversity [@Tilman2017], conservation scientists increasingly look to former agricultural lands for opportunities to restore ecosystems [@Navarro2012], sequester carbon [@Griscom2017; @Yang2020], and recover biodiversity [@Chazdon2020; @Xie2019; @Baumann2020] while incurring minimal opportunity costs. 
As people continue to move from rural areas into cities, agriculture is pushed to become more efficient, and globalization continues, some hope this trend of abandonment will continue, making such opportunities even more salient in the future [@Sanderson2018; @Taylor2021].

Yet despite substantial evidence of recent abandonment, concomitant biodiversity benefits are by no means guaranteed.
At a fundamental level, the recovery of natural ecosystems requires time, and in many ecosystems, this process can take multiple decades [@Acevedo-Charry2019; @Isbell2019; @Nerlekar2020].
If abandonment is relatively short-lived, and former fields are recleared and recultivated within a decade or two, as recent evidence indicates occurs in abandoned agricultural lands [@Crawford2022; @Dara2018; @Smaliychuk2016] and Neotropical secondary forests [@Reid2018b; @Nunes2020; @Schwartz2020; @Sloan2022], ecosystems may not have enough time to accumulate substantial gains in carbon or biodiversity.
Moreover, agricultural abandonment can have both positive and negative socioeconomic and environmental consequences, all of which can be highly dependent on climate and biophysical factors, past management history, and various human factors [@ReyBenayas2007; @Ustaoglu2018].
Given that individual species have unique biophysical and habitat requirements and are not equally distributed around the world, spatial and temporal variation in abandonment may also play a substantial role in determining outcomes for biodiversity.
Together, these factors have contributed to an ongoing controversy about the extent to which abandonment will benefit biodiversity [@Queiroz2014; @VanderZanden2017; @Barnaud2021].

This controversy is perhaps most prominent in Europe, where a substantial amount of conservation attention has been afforded to "farmland biodiversity:" species of conservation concern that are often dependent on open-habitats and low-intensity farmland, as a result of Europe's long history of cultivation, the loss of natural disturbance regimes, and cultural preferences [@Batary2015; @Fischer2008; @Fischer2012a].
When viewed from the perspective of what could be lost, abandonment is sometimes considered a threat to biodiversity, despite the potential for gains in other types of species [@Queiroz2014; @Verburg2009; @Navarro2012]. 
Variety within the natural world is one of the primary factors motivating conservation, and it should come as no surprise that different species may respond differently to land use change [@Sirami2008; @Finch2019].
But in order to better incorporate biodiversity into policy responses to abandonment, conservationists need a more nuanced understanding of which species benefit, which do not, and how these patterns balance out across whole landscapes.

Despite a large number of studies investigating the recovery of biodiversity in abandoned agricultural land over time [e.g., @Sirami2008, @Uchida2014, @Plieninger2014, @Isbell2019, @Bowen2007], few studies have attempted to combine spatially-explicit maps of abandonment with biodiversity data in order to investigate patterns in how abandonment affects biodiversity at the landscape scale.
Critically, these few studies have done so only coarsely, whether using distance to protected areas as a proxy for biodiversity value [@Meyfroidt2016], expert-designated regional priorities, coarse measures of species richness [@Henry2019], or a restricted subset of species [@VanderZanden2017].
Others have used individual species data to prioritize locations for restoration based on estimated reductions in extinction risk (e.g., @Crouzeilles2020; @Strassburg2019), but these approaches have not been linked to observed changes in land cover through time nor explored the nuances of which species benefit.
Moreover, the abandonment maps used in these analyses have historically been limited in spatial and temporal resolution, lacking the detail required to track abandonment dynamics through time.

Recent efforts to improve how abandonment is mapped from satellite imagery have produced new annual maps of cropland abandonment with higher spatial resolution (30-m resolution) and temporal coverage (1987-2017) than previous maps [@Yin2020]. 
These time series data have allowed for more detailed analysis of the long-term trajectories and persistence of abandoned cropland [@Crawford2022], and also better equip researchers to investigate the implications of abandonment for individual species across multiple landscapes and longer time horizons.

Here, we combine these new annual land cover time series for 11 sites [@Yin2020] with range maps and habitat preferences for 2023 species of birds and mammals to calculate changes in area of habitat [AOH, @Brooks2019] resulting from cropland abandonment between 1987 and 2017.
We address three questions: 
1) To what extent did cropland abandonment result in gains or losses of habitat for each of the bird and mammal species that occur at one of our sites, and how does the balance of winners and losers vary across sites and by IUCN threat status?
2) How did frequent recultivation constrain the habitat implications of abandonment, when comparing observed abandonment and recultivation with a scenario of potential abandonment in which recultivation was not allowed to occur? 
3) Finally, how do the changes resulting directly from abandonment compare to broader habitat trends that occurred during our time series, specifically when incorporating changes that occurred a) prior to cultivation and subsequent abandonment, or b) outside of abandonment, through the entire landscape at each site?

By exploring the impact of abandonment on habitat availability for every bird and mammal species that occurred at our sites, we provide a nuanced and detailed examination of how cropland abandonment affected biodiversity across a broad range of sites that have recently experienced abandonment and at multiple spatial and temporal scales.
Our analysis demonstrates that cropland abandonment holds the potential to increase habitat area for a majority of bird and mammal species, yet that these gains rarely compensate for habitat loss that took place prior to cultivation and abandonment or elsewhere at each site, in part due to the recultivation of abandoned croplands.
Together, these results highlight opportunities to improve the biodiversity outcomes of abandonment, and better harness broad changes taking place within agricultural landscapes for conservation.

# Methods {.unnumbered}

## Abandonment maps and definitions {.unnumbered}

We build on a 30-m annual land cover time series produced for eleven sites from 1987-2017 [@Yin2020] (Figs. \@ref(fig:lc-1987)) and \@ref(fig:lc-2017).
These sites, each between `r round(min(area_summary_df$total_site_area_ha_2017/10^6), digits = 2)` and `r round(max(area_summary_df$total_site_area_ha_2017/10^6), digits = 2)` million ha (Mha) in size^[For context, the state of New Jersey is about 1.90 Mha], were selected to represent areas with documented histories of cropland abandonment in the past three decades, and cover a range of biomes, including drylands, temperate regions, and the sub-tropic and wet tropics, and span four continents.
[We excluded three sites from those originally mapped by @Yin2020 that had low classification accuracy, following @Crawford2022.]
This time series mapped four land cover classes at high accuracy (average overall accuracy of $85\pm4$%): 1) cropland, 2) herbaceous vegetation (e.g., grassland), 3) woody vegetation (e.g., forests), and 4) non-vegetation (e.g., water, urban, or barren land).
Previous work [@Crawford2022] mapped abandoned croplands by identifying periods of non-cropland (grassland or woody vegetation) that followed periods of cultivation and remained non-cropland for five or more consecutive years, following a commonly used five-year abandonment definition designed to exclude short-term fallow cycles [@FAO2016].
We used these abandonment maps to identify all pixels that were abandoned at least once during the time series (the maximum extent of abandonment, which we refer to as "abandoned pixels;" Figs. \@ref(fig:abn-lc-1987)) and \@ref(fig:abn-lc-2017)).

@Crawford2022 showed that while substantial abandonment occurred at our eleven study sites (8.76 Mha of croplands were abandoned one or more times from 1987-2017), this abandonment was relatively short-lived and subject to high rates of recultivation (only 6.06 Mha of croplands remained abandoned as of 2017).
In order to assess the implications of recultivation on the impact of abandonment on biodiversity, we also use a simple scenario of "potential abandonment" developed by [@Crawford2022] that assumed that no recultivation took place during the time series. 
In this scenario, abandoned croplands were left abandoned from the year of initial abandonment through the end of the time series.

## Translating from land cover to IUCN habitat types {.unnumbered}

Because AOH is calculated based on habitat, not land cover, we translated our land cover maps into maps of habitats to match the habitat suitabilities listed by IUCN. 
To do so, we used a recently published 100-m global map of IUCN habitat types for the year 2015 developed by @Jung2020a, resampled to 30-m to match our land cover maps.
First, we developed a simple crosswalk between our coarse land cover types (woody vegetation, herbaceous vegetation, non-vegetation, and cropland) and the more specific habitat types considered by the IUCN (e.g., temperate forest, subtropical/tropical dry grassland), based on how each habitat type would likely be categorized during classification of satellite imagery into land cover (see Table \@ref(tab:iucn-crosswalk-table)).
Because this crosswalk delineates "one to many" relationships (i.e., one land cover class could be multiple IUCN habitat types), we next assigned individual pixels of land cover to the most common corresponding IUCN habitat type within the surrounding pixels in the 2015 habitat map from @Jung2020a (i.e., assigning a coarse land cover like grassland to the most common grassland habitat type within the surrounding neighborhood; see Section \@ref(si-habitat-assignment) for more details).
We considered shrubland and savanna habitat types to be a subset of our woody vegetation (i.e., forest) and herbaceous vegetation (i.e., grassland) land cover classes, respectively.
Because these habitats occupy intermediate zone between these land cover classes, we conducted a sensitivity analysis to assess similarity between land cover and reclassified habitat maps under the full range of possible categorizations (see Section \@ref(si-savanna-shrubland-jac) and Fig. \@ref(fig:jung-crosswalk-jac). 
We found similar levels of spatial overlap across all possible categorizations, so we categorized these habitats to match how they would be classified directly from satellite imagery (as noted previously).

## Species range maps and habitat preferences {.unnumbered}

Our analysis is built on the calculation of a species' Area of Habitat (AOH, *sensu* @Brooks2019), defined as the total area of a species range within a given site that falls within suitable habitat types and elevations, as derived from IUCN assessments.
We sourced species range maps for all bird and mammal species from Birdlife International [@BirdLife2021] and the IUCN [@IUCN2021] respectively, and extracted habitat and elevation preferences from IUCN assessments. 
We exclude marine species and species classified as Extinct (EX) or Extinct in the Wild (EW), and only considered areas where species were considered extant (presence) and native or reintroduced (origin), leaving 1495 bird and 528 mammal species (2023 total) with ranges overlapping with at least one of our eleven sites.
For species with range maps delineated by season, we only consider areas where species are resident, breeding, or non-breeding (we exclude passage areas for migratory species).
For species with range maps delineated by season (i.e., "resident," "breeding," and "wintering" ranges), we only considered those habitats marked as "suitable" in the relevant season associated with that portion of the range.
After filtering these species by elevation and seasonally-specific habitat preferences, the number of bird and mammal species with habitat within at least one site fell to 1688 (1035 birds and 498 mammals).

When comparing results by IUCN Red List status, we broadly defined species as "threatened" when assessed by the IUCN as "Vulnerable," "Endangered," or "Critically Endangered," and "not threatened" when assessed as "Near Threatened" or "Least Concern."
Species assessed as "Data Deficient" are excluded.
We classified species by range size based on whether a species total global range area (extent of occurrence) was above or below the global median [following @Jenkins2013].

Because species' habitat preferences listed in IUCN assessments do directly not account for habitat age, we conservatively excluded species if they required mature habitats in at least one season at each site, given that this would prevent such species from gaining habitat from changes that occurred without our 31 year time series. 
Bird species were given a manual binary classification coding whether a species was a mature forest obligate, based on habitat descriptions from IUCN assessments and individual species accounts in The Handbook of the Birds of the World [@BOW2022].
Here we defined a mature forest obligate as any species that required habitat older than 30 years (the maximum age of abandoned croplands was 30 years, given the bounds of our time series).
Neotropical bird species were independently verified by expert observer R. Alex Wiebe.
Mammals were not classified due to a lack of online access to The Handbook of the Mammals of the World [@HMW2019], and therefore all mammal species were included in our analysis.

## AOH scenarios {.unnumbered}

In order to compare changes in Area of Habitat due to cropland abandonment with those due to broader land cover changes prior to and outside of abandonment, we calculated AOH across three geographic and temporal scales: 1) for abandonment only, from first cultivation through 2017, 2) for abandonment only, from 1987 through 2017, and 3) for the entire landscape, 1987-2017.
In order to isolate the effect of cropland abandonment on habitat availability for each species, we first considered only changes in habitat stemming directly from cropland abandonment ("Scenario 1"). 
This involved calculated AOH across only pixels that were abandoned at some point during the time series, derived by constraining our time series of habitat types by the maximum extent of abandonment taken from @Crawford2022 (which we refer to as "abandoned pixels").
Furthermore, in order to isolate the effect of cropland abandonment on species' habitat, we only considered habitat provided by abandoned pixels from the year a given pixel was first cultivated through the end of the time series (i.e., "from first cultivation through abandonment").
In effect, we only start counting habitat for a pixel from when it was first cropland through when it was abandoned (and potentially recultivated), in order to understand the change from cropland to non-cropland.
We also calculated a slight variation of this first scenario in order to explore the impact of recultivation on AOH, calculating AOH in abandoned pixels for our estimate of "potential abandonment," in which no recultivation took place following abandonment ("Scenario 1P").

In order to put the AOH changes results directly from abandonment in a broader context of habitat loss and wider land use changes at each site, we widened out temporal and spatial scope for calculating AOH. 
Our second scenario ("Scenario 2") considered changes in habitat that took place in abandoned pixels relative to the start of our time series.
Like our first AOH calculation, we restricted AOH to only pixels that experienced abandonment, but now considered the complete trend in habitat across the full time series (1987-2017), in order to capture any habitat changes that occurred before cultivation and subsequent abandonment (e.g., deforestation).
Third, we calculated AOH for the entire landscape at each site, across the entire time series (1987-2017), therefore capturing the overall trend in AOH that each species experienced during our study period, including abandonment but also other land cover change dynamics ("Scenario 3"). 

## Summarizing AOH trends {.unnumbered}

We extracted trends in AOH over time for each species at each site using simple linear regressions parameterized to predict AOH as a function of year for each species at each site (converting raw AOH observations, shown in Fig. \@ref(fig:aoh-sp-obs-crop-abn), into trend lines, shown in Fig. \@ref(fig:aoh-trend-lines-crop-abn)).
AOH trends were classified as either gains or losses of habitat over time when slope coefficients were positive or negative, respectively, and statistically significant (p < 0.05, where the 95% confidence interval of the slope coefficient did not overlap with zero).
Species with AOH slopes that were not statistically significant ($p\geq0.05$) were classified as having "No Trend" at a given site. 
Species that gain habitat over time are referred to as "winner" species, while species that lose habitat are referred to as "loser" species.

Because each species could occur at multiple sites, there were some cases where a individual species had multiple types of trends at different sites.
When counting species' trends across all sites, we grouped species based on the consistency of trend direction.
Species with consistently positive trends ("Gain"), negative trends ("Loss"), or insignificant trends ("No Trend") across all sites where they occur were classified simply based on the single consistent trend. 
Species with a significant trend at one or more sites, but no significant trend at other sites, were considered to have a "weak" trend overall ("Weak Gain" or "Weak Loss", depending on the direction of the trend, where significant).
Species with significant gains at one or more sites and significant losses at others, were considered to have "Opposite" trends, where the impact of abandonment is context dependent.

## Data processing {.unnumbered}

We analyzed spatial data (maps of land cover, abandonment, IUCN habitat types, species ranges, and elevation), IUCN assessment data, and resulting area of habitat data in RStudio version `r rstudioapi::versionInfo()$version` [@RStudio], using `r R.Version()$version.string` [@R-base], using the terra [@R-terra], data.table [@R-data.table], and tidyverse [@R-tidyverse] packages.

# Results {.unnumbered}

(ref:caption-aoh-by-site-crop-abn) Number of bird and mammal species shown by trend in AOH in abandoned pixels starting from the year of first cultivation and running through 2017. Species counts are shown for individual sites and summed across all sites ("Overall" at left). Observed AOH and modeled trends for individual species are shown in Figs. \@ref(fig:aoh-sp-obs-crop-abn) and \@ref(fig:aoh-trend-lines-crop-abn).

```{r aoh-by-site-crop-abn, fig.cap = '(ref:caption-aoh-by-site-crop-abn)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_overall_w_site_combo_crop_abn_iucn",
  aoh_run_date, "_no_lab.pdf"))
```

## Changes in habitat area resulting from cropland abandonment {.unnumbered}

Overall, we found that a majority of bird and mammal species gained habitat as a result of abandonment: `r aoh_stats_main$n_gain_bird` bird and `r aoh_stats_main$n_gain_mam` mammal species showed statistically significant and positive trends in AOH over time everywhere they occurred, compared to `r aoh_stats_main$n_loss_bird` bird and `r aoh_stats_main$n_loss_mam` mammal species that lost habitat as a direct result of abandonment (Fig. \@ref(fig:aoh-by-site-crop-abn)).
When considering each species' trend at each site, `r aoh_stats_main$wl_prop_bird_site_sp` times as many birds and `r aoh_stats_main$wl_prop_mam_site_sp` times as many mammals gained habitat, on average, than lost habitat.^[*Note:* when looking at overall trends for each species, the proportion of winners to losers actually increases (to `r aoh_stats_main$wl_prop_bird_sp`x and `r aoh_stats_main$wl_prop_mam_sp`x for birds and mammals respectively for consistent winners and losers, and `r aoh_stats_main$wl_prop_bird_sp_w_weak` and `r aoh_stats_main$wl_prop_mam_sp_w_weak` when including "weak" trends.]
This pattern varied across sites, but was generally consistent, ranging from `r aoh_stats_main$wl_prop_site_min` to `r aoh_stats_main$wl_prop_site_max` times as many winners as losers at different sites. 
The lone exception were birds in Iraq (our site with the lowest species richness; Fig. \@ref(fig:richness-by-site)), where only `r aoh_stats_main$wl_iraq` times as many birds gained habitat on average.
Though bird and mammals species were each much more likely to gain habitat than to lose habitat due to abandonment, in general, a relatively greater share of bird species lost habitat than mammal species, both overall (Fig. \@ref(fig:aoh-overall-stacked)) and at individual sites (Fig. \@ref(fig:aoh-by-site-stacked)).

When considering species of conservation concern, we found that a relatively greater proportion of threatened species gained habitat as a result of abandonment (Fig. \@ref(fig:aoh-iucn-by-site-crop-abn)), though there were relatively low numbers of threatened species at each site (Fig. \@ref(fig:richness-by-site)).
This pattern also held, though a bit more weakly, across individual IUCN statuses (Fig. \@ref(fig:aoh-by-iucn)), which showed that species in higher risk categories gained relatively more habitat.
Though relatively few small-ranged birds and mammals were affected by abandonment at our sites (3 bird and 6 mammal species), we also found that nearly all small-ranged species gained habitat as a result of abandonment (with the exception of one small-ranged bird: the Iraq babbler, *Argya altirostris*; see Fig. \@ref(fig:aoh-by-range)).

(ref:caption-aoh-iucn-by-site-crop-abn) The number of species with various trends in AOH in abandoned pixels from cultivation through 2017 (Scenario 1), separated into species classified by the IUCN as "threatened" ("Vulnerable," "Endangered," or "Critically Endangered") or "not threatened" ("Near Threatened" or "Least Concern"). Species classified as "Data Deficient" are excluded. Site totals are shown on the left, and overall totals are shown on the right, with birds on the top half and mammals on the bottom half. Results for Scenarios 2 and 3 are shown in Figs. \@ref(fig:aoh-by-iucn-lumped), and individual IUCN categories in Fig. \@ref(fig:aoh-by-iucn). Site specific results for scenarios 2 and 3 corresponding to this figure are shown in Figs. \@ref(fig:aoh-iucn-by-site-max-abn) and \@ref(fig:aoh-iucn-by-site-full) respectively.

```{r aoh-iucn-by-site-crop-abn, fig.cap = '(ref:caption-aoh-iucn-by-site-crop-abn)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_IUCN_by_site_combo_crop_abn_iucn",
  aoh_run_date, ".pdf"))
```

## Effects of recultivation on habitat area {.unnumbered}

If abandoned croplands had not been recultivated during our time series (as assumed in our scenario of "potential abandonment," Scenario 1P), we found that the share of species that gain habitat would have increased for both birds and mammals overall (Fig. \@ref(fig:aoh-obs-v-pot-by-site)).
Many species with a "Weak Gain" trend or "No Trend" for observed abandonment (Scenario 1) instead experienced consistent gains across sites in our potential abandonment scenario.
Yet while the number of winning species increased, so too did the number of losing species: a greater number of species had statistically significant losses of habitat in this scenario of potential abandonment.
Most sites, however, did not see substantial differences in the numbers of winners and losers between observed and potential abandonment.
The single exception was Mato Grosso, Brazil, where `r obs_v_pot_stats$mg_n_winner_bird_pot - obs_v_pot_stats$mg_n_winner_bird_obs` more bird species and `r obs_v_pot_stats$mg_n_winner_mam_pot - obs_v_pot_stats$mg_n_winner_mam_obs` more mammal species on net showed statistically significant positive trends in AOH during our time series when recultivation was assumed not to occur, largely resulting from the transition of many species from "no trend" to "gain" (Fig. \@ref(fig:aoh-obs-v-pot-by-site)).

We also observed substantial differences in the magnitude of area of habitat gained (or lost) by the end of the time series, even for species whose overall trends did not change.
In general, trends for individual species were amplified: winner species gained (and loser species lost) relatively more habitat by the end of the time series when recultivation was not allowed to occur (Figs. \@ref(fig:aoh-p-change-obs-v-pot-bird) - \@ref(fig:aoh-p-change-obs-v-pot-mam)).
This effect was more intense for species that lost habitat: at most sites, losing species lost a relatively greater amount of habitat in the potential scenario than winning species gained, though the number of losing species remained lower than winning species everywhere except Iraq.
Species that lost habitat in both scenarios (`r obs_v_pot_stats$n_same_loss_bird` bird-site pairs and `r obs_v_pot_stats$n_same_loss_mam` mammal-site pairs) lost on average `r obs_v_pot_stats$mean_p_loss_same`% more habitat by the end of the time series in the potential scenario relative to the observed abandonment scenario (ranging between `r obs_v_pot_stats$mean_p_loss_site_min`% and `r obs_v_pot_stats$mean_p_loss_site_max`% greater losses at individual sites).
In contrast, species that gained in both scenarios (`r obs_v_pot_stats$n_same_gain_bird` bird-site pairs and `r obs_v_pot_stats$n_same_gain_mam` mammal-site pairs) only gained on average `r obs_v_pot_stats$mean_p_gain_same`% more habitat (ranging between `r obs_v_pot_stats$mean_p_gain_site_min`% and `r obs_v_pot_stats$mean_p_gain_site_max`% greater gains at individual sites).

In Mato Grosso, however, this trend reversed, affecting a large number of species given this site's high richness.
Here, `r obs_v_pot_stats$mg_improve_n_total` species improved trend (`r obs_v_pot_stats$mg_improve_n_bird_no_trend` and `r obs_v_pot_stats$mg_improve_n_mam_no_trend` birds and mammals changing from no trend to significant gains, respectively, and `r obs_v_pot_stats$mg_improve_n_bird_loss` birds and `r obs_v_pot_stats$mg_improve_n_mam_loss` mammals changing from loss to gain, respectively), gaining on average `r obs_v_pot_stats$mg_improve_mean_p_gain`% more AOH by the end of the time series in the scenario of potential abandonment.
For comparison, only `r obs_v_pot_stats$n_sp_improve_non_mg_bird` bird and `r obs_v_pot_stats$n_sp_improve_non_mg_mam` mammal species outside of Mato Grosso improved trend.
Meanwhile, `r obs_v_pot_stats$mg_worsen_n_bird` bird and `r obs_v_pot_stats$mg_worsen_n_mam` mammal species at Mato Grosso had their AOH trend worsen, and AOH declining by `r obs_v_pot_stats$mg_worsen_mean_p_loss`% on average relative to observed abandonment.

(ref:caption-aoh-obs-v-pot-by-site) The number of species with various trends in AOH for observed abandonment (Scenario 1) and our scenario of potential abandonment without recultivation (Scenario 1P). Site totals are shown on the left, and overall totals are shown on the right, with birds on the top half and mammals on the bottom half. Differences in the magnitude of AOH gained or lost are shown in Figs. \@ref(fig:aoh-p-change-obs-v-pot-bird) and \@ref(fig:aoh-p-change-obs-v-pot-mam)).

```{r aoh-obs-v-pot-by-site, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_obs_v_pot_by_site_combo",
                                      aoh_run_date,".pdf"))
```

## Abandonment in the context of broader land cover changes {.unnumbered}

When incorporating land cover changes that occurred prior to cultivation (Scenario 2) or considering the entire landscape of each site (Scenario 3), we find that the balance of winners to losers becomes much less favorable (Fig. \@ref(fig:aoh-overall)).
Whereas the majority of bird and mammal species gained habitat from cropland abandonment alone, in these two scenarios, many more species had losing trends in AOH. 
In both scenarios, there were more bird species that lost habitat than gained (`r aoh_stats_max$n_gain_bird` winners and `r aoh_stats_max$n_loss_bird` losers in Scenario 2, and `r aoh_stats_full$n_gain_bird` winners and `r aoh_stats_full$n_loss_bird` losers in Scenario 3).
Mammals had about equal numbers of winners and losers in both scenarios: slightly more mammals gained habitat in Scenario 2 (`r aoh_stats_max$n_gain_mam` winners vs. `r aoh_stats_max$n_loss_mam` losers), but more mammals lost habitat overall when incorporating land use changes outside of abandonment in each site (`r aoh_stats_full$n_gain_mam` winners and `r aoh_stats_full$n_loss_mam` losers; Scenario 3).
For both vertebrate classes, the number of species with no trend or contrasting trends across sites ("Opposites") increased, and the number of loser species did not reach a majority (Fig. \@ref(fig:aoh-overall-stacked)).

These trends again varied by site, but in general, the number of losing species increased across all sites when incorporating land cover changes that occurred prior to cultivation (Fig. \@ref(fig:aoh-by-site-max-abn); Scenario 2), or outside of abandonment (Fig. \@ref(fig:aoh-by-site-full); Scenario 3).
These trends were much stronger for birds than for mammals
At some of the sites with the clearest increases in losing bird species between Scenario 1 and Scenarios 2 and 3 (e.g., `r paste0(long_labels[c(1,8, 10)], collapse = "; ")`), the number of mammal species that gained habitat overall still remained higher than the number of species that lost habitat (Fig. \@ref(fig:aoh-by-site-stacked); see also Figs. \@ref(fig:aoh-by-site-max-abn) and \@ref(fig:aoh-by-site-full)).

Differences between threatened and non-threatened species were even greater for birds when accounting for habitat loss prior to cultivation and land cover changes outside abandoned croplands that were abandoned (Fig. \@ref(fig:aoh-by-iucn-lumped)).
In fact, a greater share of threatened birds gained habitat than lost habitat, even in Scenarios 2 and 3, in contrast to non-threatened birds.
On the other hand, a greater share of threatened mammal species gained habitat than lost habitat even when accounting for prior habitat loss prior to abandonment (Scenario 2), but not when including land cover changes across the entire landscape at each site (Scenario 3). 
Moreover, though threatened mammal species responded more favorably than non-threatened mammal species in Scenario 2, the opposite was true in Scenario 3, where a greater share of threatened species lost habitat than non-threatened species (Fig. \@ref(fig:aoh-by-iucn-lumped))
 
When considering our scenario of potential abandonment for these broader AOH scenarios, we found that the balance again flipped in favor of winner species when considering habitat losses that occurred prior to cultivation and subsequent abandonment, for both birds and mammals (Fig. \@ref(fig:aoh-overall-w-pot), Fig. \@ref(fig:aoh-overall-stacked-w-pot)).
Reducing recultivation also led to a relative increase in the number of species that gained habitat across the entire landscape (Scenario 3P), resulting in slightly more winners than losers for mammals.
Though the number of birds that gained habitat increased in this potential scenario, it remained lower than the number of losers overall, and there was wide variation in how the ratio of winners to losers changed across individual sites (\@ref(fig:aoh-by-site-full-potential)).

(ref:caption-aoh-overall) Number of bird and mammal species with various trends in AOH during our time series. Columns correspond to three scales of AOH calculation: only abandoned pixels, from first cultivation through 2017 (Scenario 1); only abandoned pixels, from 1987-2017 (Scenario 2); and the entire landscape at each site, from 1987-2017 (Scenario 3). Trends in AOH were determined through linear regression of each species AOH over time at each site. Trends that were not statistically significant (p >= 0.05) are shown as "No Trend." Species with consistently positive trends ("Gain"), negative trends ("Loss"), or insignificant trends ("No Trend") are shown in dark green, yellow, and black, respectively. Species that had a significant trend at one or more sites, and no trend at one or more other sites, are labelled with "Weak Gain" (light green) or "Weak Loss" (light yellow). Species with significant gains at some sites and significant losses at others are marked as "Opposites" (gray). Proportional results are shown in \@ref(fig:aoh-overall-stacked). Site specific results for Abandonment 1987-2017 (Scenario 2) and the Entire Landscape 1987-2017 (Scenario 3) are shown in Figs. \@ref(fig:aoh-iucn-by-site-max-abn) and \@ref(fig:aoh-iucn-by-site-full), respectively. Results for our scenarios of "potential abandonment" are shown in Figs. \@ref(fig:aoh-overall-w-pot) and \@ref(fig:aoh-overall-stacked-w-pot)

```{r aoh-overall, fig.cap = '(ref:caption-aoh-overall)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_by_vert_aoh_type", aoh_run_date,".pdf"))
```

# Discussion {.unnumbered}

When considering differences between threatened and non-threatened species,

Outline:

1. Start with a short paragraph restating the general finding of the paper. 
  - Abandonment seems to be, on the whole, positive for more species than it harms
  - But mammals seem less sensitive to abandonment, and more likely to see gains
  - There is strong variation across sites, and given different identities of species at each site, policies designed to address abandonment should 
2. Then go into what our results mean: 
  - differences between birds and mammals - In part, this means that a relatively greater share of bird species can find meaningful habitat in croplands, but it cannot be attributed to habitat generalist species alone, because these species that occcur equally well in cropland as grassland or forest would not see declines in area as a result of the replacement of cropland with grasslands or regenerating forests. However,  
  -
3. Results in context of previous studies.
  - @Queiroz2014, framing matters. 
  - Our results contrast with the prevailing view in western Europe that much biodiversity may be threatened by abandonment. However, the nuances of our site-by-site results make clear that each region should be considered separately. -specific  indicate that this might be a som, where evidence points to abandonment a greater share of 
3. Limitations
  - of abandonment data: Cautioning against putting too much stock in Mato Grosso, given the relatively little abandonment that took place there. However, the degree to which species trends improved it is a testament to the power of secondary habitat regeneration at this site,...
  - of biodiversity data, regeneration: successful regeneration is not a guarantee, even if protected from recultivation. And,  
4. Areas for future exploration. 
  - Perhaps looking at population trends over time.
  - On the whole, our results are encouraging that the majority of species gain, on the whole, as a result of abandonment. Even more importantly, our results indicate that threatened species may be even more likely than non-threatened species to gain habitat through abandonment.
  - However, losing species should be treated with care, and when necessary, incorporated into conservation planning to ameliorate habitat losses.
  - We excluded amphibians and reptiles from our study due to a relative lack of information about their distributions, habitat requirements, and threat status. However, as recent efforts to comprehensively assess reptiles reach their conclusion [@Cox2022], conservation scientists should be better equipped to incorporate these species into future analyses. 
5. Conclusions and conservation implications





Our comparison of observed and potential abandonment indicated that a large number of species at Mato Grosso stood to benefit from the reduction of recultivation, as evidenced by the transition of 82 bird

Though Mato Grosso showed strong trends in the number of winners and losers, we emphasize that this site experienced the least abandonment overall, and these patterns should be placed in appropriate context relative to other sites, where species gained (or lost) much greater amounts of habitat as a result of abandonment.


Visual inspection of the raw AOH curves for individual species 

Some strange results about area differences in the magnitude of area changes.
If our results hold, it may indicate that even if a greater proportion of species benefit from abandonment, those species that lose habitat may require special action to address the steep declines of habitat resulting from abandonment.
This is particularly the case for the development of any programs to reduce recultivation.

## Abandonment in context {-}

Finally, discuss cropland abandonment in the context of the full time series, and the full landscape.

Cropland abandonment *per se* compared with trends in abandoned pixels for 1987-2017, and compared to the full landscape (1987-2017).
Our initial results on cropland abandonment *per se* tell us whether abandonment of cropland causes a gain or loss of habitat for each species. 
By looking at the trend for those same abandoned pixels, but starting from the beginning of our time series (1987), puts gains from abandonment in the context of any habitat loss that occurred prior to cultivation.
Taking this one step further, by assessing trends of AOH for the full site area across the full time series, we put our results in context of broader trends taking place at each site. 
For discussion:

(Maybe a separate point, or moved to a later section about the different scales of AOH.)
This pattern is reversed for birds, however, when considering the full time series, even when only considering abandoned pixels.
This is likely because many species might lose habitat during conversion to cropland, but not gain it entirely back as a result of abandonment, whether due to recultivation or a lack of time for succession. 
This pattern is even stronger for mature forest obligate bird species (Fig. \@ref(fig:aoh-overall-mature-for-obl)). 
Even when excluding these species, the balance of winners to losers still shows more losers than winners for both abandonment (1987-2017) and the full landscape (1987-2017).


## Study limitations {.unnumbered}

Outline:

- Fragmentation
- Habitat preferences as a function of age of habitat, succession. Regeneration can be successful, but it typically takes a long time (> 50 years) for regenerating habitats to recover levels of species richness, and even longer to recover to similar community compositions.
- Study site size: because our land cover time series was restricted to footprint of our sites, we were unable to assess how habitat gains or losses affected extinction risk for species, without an idea of changes taking place elsewhere in each species' range. 
- A lack of disturbance regimes, and incomplete regeneration on abandoned land. 

Locations
- Despite our positive results, concerns about the potential for abandonment to negatively impact species of conservation concern in some locations merit thoughtful consideration of how abandonment is likely to affect biodiversity in different sites. 

## Policy implications {.unnumbered}


## Conclusions {-}

-	Abandonment *per se* benefitted the majority of bird and mammal species at our sites, though the degree to which this was the case varied across sites (Iraq was the lone exception).
-	Threatened species were even more likely to be winner species than non-threatened species.
-	When recultivation was not allowed to take place, these trends intensified, and the number of species without significant trends declined, leading to more winners and more losers. This effect was relatively muted at many places, though Mato Grosso saw a huge increase in winners.
-	Despite many species not changing trend direction, recultivation did result in less habitat for each species overall by the end of the time series The magnitude of area gained by winner species increased by 12% on average. However, losing species lost a relatively greater share of habitat, however, losing about 72% more habitat. 
-	Though abandonment overwhelmingly led to gains in habitat for the majority of species, these gains did not compensate for habitat loss that occurred prior to cultivation (and subsequent abandonment) or elsewhere in entire landscape at each site. The plurality of birds had significantly declining trends in habitat in both of these scenarios, and while mammals were less affected, they still saw more losers than winners when incorporating the entire landscape for each site.
-	Importantly, if recultivation was not allowed to take place, abandonment would again benefit more species than it harmed, even when factoring in habitat loss prior to cultivation.


[This must include a statement of the key conclusions of the work, and implications for conservation policy and practice.]

Other research has indicated that when possible, the environmental impacts of agriculture (in terms of rarity-weighted species richness, carbon stocks, and water use) could be lessened through relocation and regeneration [@Beyer2022].
Our results provide evidence that, on a multi-decadal timescale, cropland abandonment can result in gains in habitat for a majority of individual species.
Considering each species individually is critical from a conservation perspective, in order to account for differences in species responses. 
However, our results indicate that concerns of abandonment overwhelmingly harming species of conservation concern seem not to hold at our sites.

Yet despite how encouraging these results are, our additional analyses emphasize that there is significantly more that can and should be done to maximize the conservation potential held by abandoned agricultural lands, particularly by providing incentives to reduce the recultivation of abandoned croplands. 
Our results also demonstrate that even in regions with high levels of documented abandonment, where one could expect abandonment to have the greatest potential to benefit biodiversity (if those benefits are realized), we see that too often, these gains do not outweigh losses that took place either prior to or outside of abandonment.
Reducing habitat loss remains a critical tool for conservation, even while simultaneously working to maximize the returns from broad changes taking place in agricultural landscapes. 






# Acknowledgements (100 words maximum) {.unnumbered}

We thank the following researchers at the University of Wisconsin-Madison, who generously shared the land cover maps that made our analysis possible: Amintas Brando, Johanna Buchner, David Helmers, Benjamin G. Iuliano, Niwaeli E. Kimambo, Katarzyna E. Lewiska, Elena Razenkova, Afag Rizayeva, Natalia Rogova, Seth A. Spawn-Lee, and Yanhua Xie.
We thank the Drongos research group for advice and companionship.
Our analysis would not be possible without the dedication of IUCN, BirdLife International, and numerous individual experts who contributed to species assessments.
Analyses were performed using Princeton Research Computing resources at Princeton University.
This research was supported by the High Meadows Foundation (DSW) and the NASA Land Cover and Land Use Change Program (Grant no. 80NSSC18K0343, to VCR, HY), and contributes to the Global Land Programme (<https://glp.earth/>).

## Author contributions {.unnumbered}

TBD

<!-- CLC and DSW conceived of the project in collaboration with HY and VCR. -->
<!-- HY contributed guidance on data curation and the analysis of raw land cover data. -->
<!-- CLC designed and conducted the data analysis, with feedback from HY, DSW, and VCR. -->
<!-- All authors contributed to writing the manuscript, with CLC preparing the initial draft. -->

## Competing interests {.unnumbered}

The authors declare no competing interests.

## Data and code availability {.unnumbered}

The annual land cover maps [@Yin2020] and abandonment maps [@Crawford2022] that provide the foundation for our analysis are archived at Zenodo (<https://doi.org/10.5281/zenodo.5348287>). 
Code to replicate these analyses is available on GitHub at <https://github.com/chriscra/biodiversity_abandonment>.
Bird and mammal range maps are available upon request from BirdLife International (<http://datazone.birdlife.org/species/requestdis>) and IUCN (<https://www.iucnredlist.org/resources/spatial-data-download>) respectively.^[Range maps downloaded on January 4th, 2022.]
Species assessment data (including habitat and elevation preferences) are freely available from IUCN (<https://www.iucnredlist.org/>) 
The 2015 global map of IUCN habitat types developed by @Jung2020a is freely available at <https://zenodo.org/record/4058819>.

# References {.unnumbered}

::: {#refs}
:::

\newpage

\beginsupplement

# Supplementary information {.unnumbered}

<!-- \tableofcontents -->
<!-- \listoftables -->
<!-- \listoffigures -->

<!-- \newpage -->

# Extended Methods

## Assigning land cover to habitat types {#si-habitat-assignment}

We developed a method for assigning pixels in our coarse land cover classes to finer habitat types to match the habitats used by the IUCN to describe species' habitat preferences.
This process was based on a 2015 map of IUCN habitat types at 100-m resolution produced by @Jung2020a based on a dynamic crosswalk between the Copernicus land cover maps and the IUCN habitat types.
We in turn developed a simple crosswalk between our coarse land cover types and the finer IUCN habitat types mapped by @Jung2020a, as shown in Table \@ref(tab:iucn-crosswalk-table). 
But because there was not perfect spatial overlap between our land cover maps and the 2015 habitat map, we developed a method for assigning our pixels of land cover to the nearest habitat type on the @Jung2020a map that matched the broad category according to our crosswalk. 
We first extracted only those pixels with habitat types that fell into each of our four land cover classes, and used a focal filter to assign habitat types to empty pixels based on the most common (i.e., modal) habitat type within the surrounding nine-pixel neighborhood (using the function `terra::focal(fun = "modal")`).
This produced four maps of potential habitat types for each pixel corresponding to each coarse land cover type.
We then masked each of these potential habitat types maps by the corresponding land cover pixels in each year at each site, producing maps of the likely habitat type for each pixel of land cover, in each year.
We then combined these four back together again to produce a map of habitat types at each site in each year, which served as input for our calculation of area of habitat (AOH). 

See maps below.

### Allocating savanna and shrubland habitats to land cover classes {#si-savanna-shrubland-jac}

Allocating the finer scale IUCN habitat types to our coarse land cover classes was straightforward for forests, grasslands, non-vegetation habitat types, and arable land. 
It was less straightforward for savanna and shrubland habitats, which are intermediate between entirely open grasslands and closed forests.

Sites with significant area classified as shrubland include:
- Belarus (14%)
- Bosnia & Herzegovina (7%)
- Wisconsin (9%)
- Iraq, Nebraska, and Shaanxi (1-2%)
- Chongqing, Volgograd, and Orenburg (~0.5%)

Sites with significant area in savanna include:
- Goias (69%)
- Chongqing (19%)
- Mato Grosso (10%)

Our default was to classify savanna and shrubland habitats as grassland and forest land cover respectively, but we explored the consequences of alternative classifications by testing spatial similarity between the resulting reclassified maps. 

We calculated Jaccard similarity between our @Yin2020 land cover maps (2015) and @Jung2020a habitat maps (2015) when reclassifying savanna and shrubland habitats as either forest or grassland.
The Jaccard similarity index is a measure of spatial similarity (i.e., overlap) between two sets, defined as the proportion of shared elements, or the intersection divided by the union (Equation \@ref(eq:jaccard-equation)) [@Legendre2012]:

\begin{equation}
J(a,b) = \frac{a\cap b}{a\cup b} (\#eq:jaccard-equation)
\end{equation}

We tested the full range of reclassifications, and the Jaccard similarity between the reclassified maps are shown in Fig. \@ref(fig:jung-crosswalk-jac).

We found that:

1. At the three sites with significant amounts of shrublands (Belarus, Bosnia & Herzegovina, and Wisconsin) - which are all in forested biomes - the similarity between grassland maps increases when shrublands are considered grasslands (as do forest maps). The gold and black bars look basically the same for these three sites, which means that the increase is due to the shrub -> grassland crosswalk; the crosswalk for savanna does not seem to matter, likely because there is very little savanna at these sites. This pattern also holds for Iraq, which had about 2% shrubland (compared to 49% desert, 42% cropland, 4% grassland, and 2% urban).

2. Sites with significant area in savanna (Goias, Mato Grosso, and Chongqing, though I know this isnt really savanna) show greater similarity when savanna is classed as grassland, not as forest (this makes sense, given that these sites likely have a greater distinction between true forest and savanna). The green and gold bars are the same, signifying that the shrubland decision doesnt matter much (because there is so little shrubland there).

3. The four remaining sites - Nebraska, Orenburg, Shaanxi, and Volgograd - are all in grassland/steppe biomes, where the majority of the site is classified as grassland. At these places, the similarity is higher when shrubland is classified as forest (shown by the green and gray bars being higher).  These increases are pretty small, however, and there is relatively little shrubland at any of these sites (max. 1.4% at Nebraska)

My conclusions from this are that, across the board, I should classify savanna as grassland. For shrubland, however, because there is variation in how similar land cover maps are under these different reclassification rules, but no clear trend, we consider shrubland habitats to be a subset of forest land cover, given that shrublands are most similar spectrally to forests, from a remote sensing perspective. 

(ref:caption-jung-crosswalk-jac) Jaccard Similarity between @Yin2020 land cover maps (2015) and @Jung2020a habitat maps (2015) when reclassifying savanna and shrubland habitats as either forest or grassland.

```{r jung-crosswalk-jac, fig.cap = '(ref:caption-jung-crosswalk-jac)'}
knitr::include_graphics(path = paste0(p_plots, "crosswalk_jaccard.pdf"))
```

\newpage

# Supplemental Tables

```{r iucn-crosswalk-table, echo = FALSE, message = FALSE, eval = TRUE}
iucn_crosswalk %>%
  left_join(tibble(lc = 1:4, 
                   lc_name = c("Non-vegetation", "Woody veg. (Forest)",
                               "Cropland", "Herbaceous veg. (Grassland)"))
            ) %>%
  select(lc, lc_name, IUCNLevel, map_code) %>%
  arrange(lc) %>% #print(n = 50)

  knitr::kable(., 
             col.names = c("Land cover code", 
                           "Land cover class", 
                           "IUCN habitat",
                           "IUCN map code"),
             align = "r", 
             #format = "simple",
             format = my_format,
             booktabs = TRUE,
             longtable = TRUE,
             caption = "Crosswalk between land cover classes in Yin et al. 2020 and IUCN habitat types mapped by Jung et al. 2020."
             ) %>%
  kableExtra::kable_styling(
    # latex_options = "scale_down"
    latex_options = c("hold_position", "repeat_header")) %>%
  # row_spec(11, hline_after = TRUE) %>%
  # row_spec(12, bold = TRUE) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2, width = "4.5cm") %>%
  column_spec(3, width = "7cm") %>%
  column_spec(4, width = "2cm")
```

# Supplemental Figures

(ref:caption-richness-by-site) Species richness at each site, or the count of species with some Area of Habitat (AOH) at each site. Richness is shown for threatened ("Vulnerable," "Endangered," or "Critically Endangered") or non-threatened ("Least Concern", "Near Threatened"), as assessed by the IUCN. Species classified as "Data Deficient" are excluded.

```{r richness-by-site, fig.cap = '(ref:caption-richness-by-site)'}
knitr::include_graphics(path = paste0(
  p_plots, "richness_by_site_w_category_facet_color.pdf"))
```

## Maps

Include:

1. Habitat types of abandonment at each site in five-year periods
2. Habitat maps for 2015, directly from Jung et al. 2021.
3. Land cover of abandonment (perhaps for each five-year period)
4. Land cover of whole landscape (for five-year periods)

(ref:caption-abn-lc-1987) Land cover as of the year 1987, shown for all pixels that were abandoned at least once during the time series.

```{r abn-lc-1987, fig.cap = '(ref:caption-abn-lc-1987)'}
knitr::include_graphics(path = paste0(p_plots, "abn_land_cover_1987.pdf"))
```

(ref:caption-abn-lc-2017) Land cover as of the year 2017, shown for all pixels that were abandoned at least once during the time series.

```{r abn-lc-2017, fig.cap = '(ref:caption-abn-lc-2017)'}
knitr::include_graphics(path = paste0(p_plots, "abn_land_cover_2017.pdf"))
```

(ref:caption-lc-1987) Land cover as of the year 1987, shown for all pixels at each site.

```{r lc-1987, fig.cap = '(ref:caption-lc-1987)'}
knitr::include_graphics(path = paste0(p_plots, "land_cover_1987.pdf"))
```

(ref:caption-lc-2017) Land cover as of the year 2017, shown for all pixels at each site.

```{r lc-2017, fig.cap = '(ref:caption-lc-2017)'}
knitr::include_graphics(path = paste0(p_plots, "land_cover_2017.pdf"))
```

## Area line charts:

Two types of area

Three AOH types:

## AOH Results

### Observed AOH and modeled AOH trends

(ref:caption-aoh-sp-obs-crop-abn) Observed AOH for all bird and mammals species occurring at each site. AOH is calculated for only pixels that were abandoned at least once, from the year of first cultivation through 2017 (Scenario 1). Each individual species' trend is visualized with moderate transparency, so that darker trend lines indicate many species with overlapping trends.

```{r aoh-sp-obs-crop-abn, fig.cap = '(ref:caption-aoh-sp-obs-crop-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_sp_obs_by_site_crop_abn_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-trend-lines-crop-abn) AOH trends for all bird and mammals species occurring at each site. AOH is calculated for only pixels that were abandoned at least once, from the year of first cultivation through 2017 (Scenario 1).

```{r aoh-trend-lines-crop-abn, fig.cap = '(ref:caption-aoh-trend-lines-crop-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_crop_abn_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-sp-obs-crop-abn-potential) AOH over time for all bird and mammals species occurring at each site, for our scenario of potential abandonment (Scenario 1P). AOH is calculated for only pixels that were abandoned at least once, from the year of first cultivation through 2017, but under the assumption that no recultivation takes place and abandonment persists from the year of initial abandonment through 2017.

```{r aoh-sp-obs-crop-abn-potential, fig.cap = '(ref:caption-aoh-sp-obs-crop-abn-potential)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_sp_obs_by_site_crop_abn_potential_iucn",
                                      aoh_run_date,".pdf"))
```

<!-- (ref:caption-aoh-trend-lines-crop-abn-potential) AOH trends for all species occurring at each site, for only abandoned pixels, from cultivation through 2017, for a scenario of potential abandonment (Scenario 1P). -->

<!-- ```{r aoh-trend-lines-crop-abn-potential, fig.cap = '(ref:caption-aoh-trend-lines-crop-abn-potential)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_crop_abn_potential_iucn", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->

(ref:caption-aoh-sp-obs-max-abn) Observed AOH over time for all bird and mammals species occurring at each site (Scenario 2). AOH is calculated for only pixels that were abandoned at least once, including habitat changes resulting from land cover changes occurring prior to cultivation (i.e., for the full time series, 1987-2017).

```{r aoh-sp-obs-max-abn, fig.cap = '(ref:caption-aoh-sp-obs-max-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_sp_obs_by_site_max_abn_iucn",
                                      aoh_run_date,".pdf"))
```

<!-- (ref:caption-aoh-trend-lines-max-abn) AOH trends for all species occurring at each site, for only abandoned pixels, from 1987-2017 (Scenario 2). -->

<!-- ```{r aoh-trend-lines-max-abn, fig.cap = '(ref:caption-aoh-trend-lines-max-abn)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_max_abn_iucn", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->

(ref:caption-aoh-sp-obs-max-abn-potential) Observed AOH over time for all bird and mammals species occurring at each site, for our scenario of potential abandonment (Scenario 2P). AOH is calculated for only pixels that were abandoned at least once, including habitat changes resulting from land cover changes occurring prior to cultivation (i.e., for the full time series, 1987-2017).

```{r aoh-sp-obs-max-abn-potential, fig.cap = '(ref:caption-aoh-sp-obs-max-abn-potential)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_sp_obs_by_site_max_potential_abn_iucn",
                                      aoh_run_date,".pdf"))
```

<!-- (ref:caption-aoh-trend-lines-max-abn-potential) AOH trends for all species occurring at each site, for only abandoned pixels, from 1987-2017, for our scenario of potential abandonment (Scenario 2P). -->

<!-- ```{r aoh-trend-lines-max-abn-potential, fig.cap = '(ref:caption-aoh-trend-lines-max-abn-potential)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_max_potential_abn_iucn", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->

(ref:caption-aoh-sp-obs-full) Observed AOH over time for all bird and mammals species occurring at each site (Scenario 3). AOH is calculated for all pixels in the entire landscape were abandoned at least once, from the year of first cultivation through 2017.

```{r aoh-sp-obs-full, fig.cap = '(ref:caption-aoh-sp-obs-full)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_sp_obs_by_site_full_iucn",
                                      aoh_run_date,".pdf"))
```

<!-- (ref:caption-aoh-trend-lines-full) AOH trends for all species occurring at each site, for the full landscape at each site, 1987-2017 (Scenario 3). -->

<!-- ```{r aoh-trend-lines-full, fig.cap = '(ref:caption-aoh-trend-lines-full)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_full_iucn", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->

(ref:caption-aoh-sp-obs-full-potential) Observed AOH over time for all bird and mammals species occurring at each site, for our scenario of potential abandonment (Scenario 3P). AOH is calculated for all pixels in the entire landscape were abandoned at least once, from the year of first cultivation through 2017.

```{r aoh-sp-obs-full-potential, fig.cap = '(ref:caption-aoh-sp-obs-full-potential)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_sp_obs_by_site_full_potential_iucn",
                                      aoh_run_date,".pdf"))
```

<!-- (ref:caption-aoh-trend-lines-full-potential) AOH trends for all species occurring at each site, for the full landscape at each site, 1987-2017, for our scenario of potential abandonment (Scenario 3P). -->

<!-- ```{r aoh-trend-lines-full-potential, fig.cap = '(ref:caption-aoh-trend-lines-full-potential)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/trend_lines_by_site_full_potential_iucn", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->



### Overall

(ref:caption-aoh-overall-stacked) Overall proportion of trend types across sites, for birds and mammals, for three AOH types (cropland abandonment only, abandonment 1987-2017, and full landscape 1987-2017).

```{r aoh-overall-stacked, fig.cap = '(ref:caption-aoh-overall-stacked)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_stacked", aoh_run_date,".pdf"))
```

(ref:caption-aoh-overall-w-pot) Number of bird and mammal species with various trends in AOH during our time series, including scenarios of potential abandonment. Columns correspond to three scales of AOH calculation, for observed and potential abandonment: only abandoned pixels, from first cultivation through 2017 (observed, Scenario 1, and potential, Scenario 1P); only abandoned pixels, from 1987-2017 (observed, Scenario 2, and potential, Scenario 2P); and the entire landscape at each site, from 1987-2017 (Scenario 3). Trends in AOH were determined through linear regression of each species AOH over time at each site. Trends that were not statistically significant (p >= 0.05) are shown as "No Trend." Species with consistently positive trends ("Gain"), negative trends ("Loss"), or insignificant trends ("No Trend") are shown in dark green, yellow, and black, respectively. Species that had a significant trend at one or more sites, and no trend at one or more other sites, are labelled with "Weak Gain" (light green) or "Weak Loss" (light yellow). Species with significant gains at some sites and significant losses at others are marked as "Opposites" (gray). Site specific results for potential abandonment for Scenarios 1, 2, and 3 are shown in Figs. \@ref(fig:aoh-by-site-crop-abn-potential), \@ref(fig:aoh-by-site-max-abn-potential), and \@ref(fig:aoh-by-site-full-potential), respectively.

```{r aoh-overall-w-pot, fig.cap = '(ref:caption-aoh-overall-w-pot)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_overall_by_aoh_type_w_pot", aoh_run_date,".pdf"))
```

(ref:caption-aoh-overall-stacked-w-pot) Overall proportion of trend types across sites, for birds and mammals, for three AOH types (cropland abandonment only, abandonment 1987-2017, and full landscape 1987-2017), for both observed and potential abandonment.

```{r aoh-overall-stacked-w-pot, fig.cap = '(ref:caption-aoh-overall-stacked-w-pot)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_trend_overall_by_aoh_type_w_pot_stacked",
  aoh_run_date,".pdf"))
```

(ref:caption-aoh-overall-mature-for-obl) Number of species with various trend types across all sites, shown for mature forest obligate and non-obligate birds.

```{r aoh-overall-mature-for-obl, fig.cap = '(ref:caption-aoh-overall-mature-for-obl)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_trend_overall_by_mature_obl",
  aoh_run_date,".pdf"))
```

<!-- (ref:caption-aoh-overall-mature-for-obl-stacked) Proportion of trend types across all sites, shown for mature forest obligate and non-obligate birds. -->

<!-- ```{r aoh-overall-mature-for-obl-stacked, fig.cap = '(ref:caption-aoh-overall-mature-for-obl-stacked)'} -->
<!-- knitr::include_graphics(path = paste0( -->
<!--   p_plots, "aoh/aoh_trend_overall_stacked_by_mature_obl", -->
<!--   aoh_run_date,".pdf")) -->
<!-- ``` -->

### By site

(ref:caption-aoh-by-site-stacked) Proportion of species at each site with trends in AOH, shown across three scales of AOH calculation.

```{r aoh-by-site-stacked, fig.cap = '(ref:caption-aoh-by-site-stacked)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/trend_summary_by_site_stacked_3",
  aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-crop-abn-potential) Number of species that gain or lose AOH over time across sites, where AOH is calculated in only those pixels that were abandoned during the time series, calculated from the first year of cultivation through 2017, for our scenario of "potential abandonment," assuming no recultivation takes place (Scenario 1P). Observed AOH for individual species are shown in Fig. \@ref(fig:aoh-sp-obs-crop-abn-potential).

```{r aoh-by-site-crop-abn-potential, fig.cap = '(ref:caption-aoh-by-site-crop-abn-potential)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_overall_w_site_combo_crop_abn_potential_iucn",
  aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-max-abn) Number of species that gain or lose AOH over time across sites, where AOH is calculated in only those pixels that were abandoned during the time series, calculated for 1987-2017 (Scenario 2). Observed AOH for individual species are shown in Fig. \@ref(fig:aoh-sp-obs-max-abn).

```{r aoh-by-site-max-abn, fig.cap = '(ref:caption-aoh-by-site-max-abn)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_overall_w_site_combo_max_abn_iucn",
  aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-max-abn-potential) Number of species that gain or lose AOH over time across sites, where AOH is calculated in only those pixels that were abandoned during the time series, calculated for 1987-2017, for our scenario of "potential abandonment," assuming no recultivation takes place (Scenario 2P). Observed AOH for individual species are shown in Fig. \@ref(fig:aoh-sp-obs-max-abn-potential).

```{r aoh-by-site-max-abn-potential, fig.cap = '(ref:caption-aoh-by-site-max-abn-potential)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_overall_w_site_combo_max_potential_abn_iucn",
  aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-full) Number of species that gain or lose AOH over time across sites, where AOH is calculated for full site area, from 1987-2017 (Scenario 3). Observed AOH for individual species are shown in Figs. \@ref(fig:aoh-sp-obs-full).

```{r aoh-by-site-full, fig.cap = '(ref:caption-aoh-by-site-full)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_overall_w_site_combo_full_iucn",
  aoh_run_date,".pdf"))
```


(ref:caption-aoh-by-site-full-potential) Number of species that gain or lose AOH over time across sites, where AOH is calculated for full site area, from 1987-2017, for our scenario of "potential abandonment," assuming no recultivation takes place (Scenario 3P). Observed AOH for individual species are shown in Fig. \@ref(fig:aoh-sp-obs-full-potential).

```{r aoh-by-site-full-potential, fig.cap = '(ref:caption-aoh-by-site-full-potential)'}
knitr::include_graphics(path = paste0(
  p_plots, "aoh/aoh_overall_w_site_combo_full_potential_iucn",
  aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-site-mature-for-obl) Proportion of species with various trends in AOH at each site, shown for mature forest obligate and non-obligate birds, across three scales of AOH calculation.

```{r aoh-by-site-mature-for-obl, fig.cap = '(ref:caption-aoh-by-site-mature-for-obl)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_site_stacked_no_mat_obl",
                                      aoh_run_date,".pdf"))
```

### By IUCN

(ref:caption-aoh-by-iucn-lumped) Proportion of species with various trends in AOH, shown for species classified by the IUCN as "threatened" ("Vulnerable," "Endangered," or "Critically Endangered") or "not threatened" ("Near Threatened" or "Least Concern"). Species classified as "Data Deficient" are excluded.

```{r aoh-by-iucn-lumped, fig.cap = '(ref:caption-aoh-by-iucn-lumped)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_IUCN_status_lumped",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-by-iucn) AOH trends, in terms of the proportion of species with a given trend direction, shown by IUCN Red List Status, across our three AOH calculations.

```{r aoh-by-iucn, fig.cap = '(ref:caption-aoh-by-iucn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_IUCN_status",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-iucn-by-site-max-abn) The number of species with various trends in AOH, calculated in abandoned pixels from 1987-2017 (Scenario 2), separated into species classified by the IUCN as "threatened" ("Vulnerable," "Endangered," or "Critically Endangered") or "not threatened" ("Near Threatened" or "Least Concern"). Species classified as "Data Deficient" are excluded. Site totals are shown on the left, and overall totals are shown on the right, with birds on the top half and mammals on the bottom half.

```{r aoh-iucn-by-site-max-abn, fig.cap = '(ref:caption-aoh-iucn-by-site-max-abn)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_IUCN_by_site_combo_max_abn_iucn",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-iucn-by-site-full) The number of species with various trends in AOH, calculated for the entire landscape at each site from 1987-2017 (Scenario 3), separated into species classified by the IUCN as "threatened" ("Vulnerable," "Endangered," or "Critically Endangered") or "not threatened" ("Near Threatened" or "Least Concern"). Species classified as "Data Deficient" are excluded. Site totals are shown on the left, and overall totals are shown on the right, with birds on the top half and mammals on the bottom half.

```{r aoh-iucn-by-site-full, fig.cap = '(ref:caption-aoh-iucn-by-site-full)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_IUCN_by_site_combo_full_iucn",
                                      aoh_run_date,".pdf"))
```

<!-- The four following figures were obviated by combo figures. -->

<!-- (ref:caption-aoh-by-iucn-lumped-numeric) (Alt2) Proportion of species with various trends in AOH, shown for species classified by the IUCN as "threatened" ("Vulnerable," "Endangered," or "Critically Endangered") or "not threatened" ("Near Threatened" or "Least Concern"). Species classified as "Data Deficient" are excluded. -->

<!-- ```{r aoh-by-iucn-lumped-numeric, fig.cap = '(ref:caption-aoh-by-iucn-lumped-numeric)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_IUCN_status_lumped_numeric", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->

<!-- (ref:caption-aoh-by-site-threatened-crop-abn) AOH trends shown at each site for only those species classified as "threatened" by the IUCN ("Vulnerable," "Endangered," or "Critically Endangered"). Shown for only pixels that were abandoned at least once during the time series, from cultivation through the end of the time series. -->

<!-- ```{r aoh-by-site-threatened-crop-abn, fig.cap = '(ref:caption-aoh-by-site-threatened-crop-abn)'} -->
<!-- knitr::include_graphics(path = paste0( -->
<!--   p_plots, "aoh/trend_summary_by_IUCN_status_by_site_crop_abn_iucn_Threatened", -->
<!--   aoh_run_date,".pdf")) -->
<!-- ``` -->

<!-- (ref:caption-aoh-by-site-threatened-max-abn) AOH trends shown at each site for only those species classified as "threatened" by the IUCN ("Vulnerable," "Endangered," or "Critically Endangered"). Shown for only pixels that were abandoned at least once during the time series, from 1987-2017. -->

<!-- ```{r aoh-by-site-threatened-max-abn, fig.cap = '(ref:caption-aoh-by-site-threatened-max-abn)'} -->
<!-- knitr::include_graphics(path = paste0( -->
<!--   p_plots, "aoh/trend_summary_by_IUCN_status_by_site_max_abn_iucn_Threatened", -->
<!--   aoh_run_date,".pdf")) -->
<!-- ``` -->

<!-- (ref:caption-aoh-by-site-threatened-full) AOH trends shown at each site for only those species classified as "threatened" by the IUCN ("Vulnerable," "Endangered," or "Critically Endangered"). Shown for AOH calculated across the full landscape at each site, from 1987-2017 -->

<!-- ```{r aoh-by-site-threatened-full, fig.cap = '(ref:caption-aoh-by-site-threatened-full)'} -->
<!-- knitr::include_graphics(path = paste0( -->
<!--   p_plots, "aoh/trend_summary_by_IUCN_status_by_site_full_iucn_Threatened", -->
<!--   aoh_run_date,".pdf")) -->
<!-- ``` -->

### Observed vs. Potential Abandonment

(ref:caption-aoh-obs-v-pot-overall) Proportion of species that gain or lose AOH as a result of abandonment during our time series, for both observed abandonment and a scenario of potential abandonment, assuming no recultivation.

```{r aoh-obs-v-pot-overall, fig.cap = '(ref:caption-aoh-obs-v-pot-overall)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_obs_v_pot_overall",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-p-change-obs-v-pot-bird) Percent change in area of habitat at the end of the time series between observed and potential abandonment, for bird species only. Relative change in trend between scenarios is show by colors, and number of species in each group is shown in numeric annotation on bars. (See Fig. \@ref(fig:aoh-p-change-obs-v-pot-mam) for a corresponding figure for mammals.)

```{r aoh-p-change-obs-v-pot-bird, fig.cap = '(ref:caption-aoh-p-change-obs-v-pot-bird)'}
knitr::include_graphics(path = paste0(p_plots,
                                      "aoh/aoh_p_change_obs_v_pot_bird",
                                      aoh_run_date,".pdf"))
```

(ref:caption-aoh-p-change-obs-v-pot-mam) Percent change in area of habitat at the end of the time series between observed and potential abandonment, for bird species only. Relative change in trend between scenarios is show by colors, and number of species in each group is shown in numeric annotation on bars. (See Fig. \@ref(fig:aoh-p-change-obs-v-pot-bird) for a corresponding figure for birds.)

```{r aoh-p-change-obs-v-pot-mam, fig.cap = '(ref:caption-aoh-p-change-obs-v-pot-mam)'}
knitr::include_graphics(path = paste0(p_plots,
                                      "aoh/aoh_p_change_obs_v_pot_mam",
                                      aoh_run_date,".pdf"))
```

<!-- Obviated by the combo figure above -->
<!-- (ref:caption-aoh-obs-v-pot-by-site-bird) Number of bird species with various trends in AOH for scenarios of observed vs. potential abandonment, shown for each site. Mature forest obligate birds are excluded. -->

<!-- ```{r aoh-obs-v-pot-by-site-bird, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site-bird)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_obs_v_pot_by_site_sbs_bird", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->

<!-- (ref:caption-aoh-obs-v-pot-by-site-mam) Number of mammals species with various trends in AOH for scenarios of observed vs. potential abandonment, shown for each site. -->

<!-- ```{r aoh-obs-v-pot-by-site-mam, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site-mam)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/aoh_trend_obs_v_pot_by_site_sbs_mam", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->

<!-- (ref:caption-aoh-obs-v-pot-by-site-stacked) Proportion of species with various AOH trends for scenarios of observed vs. potential abandonment, shown for each site -->

<!-- ```{r aoh-obs-v-pot-by-site-stacked, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site-stacked)'} -->
<!-- knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_obs_v_pot_by_site", -->
<!--                                       aoh_run_date,".pdf")) -->
<!-- ``` -->

(ref:caption-aoh-obs-v-pot-by-site-mature-for-obl) AOH for observed vs. potential abandonment, by site and AOH scale, shown for mature forest obligate and non-obligate birds.

```{r aoh-obs-v-pot-by-site-mature-for-obl, fig.cap = '(ref:caption-aoh-obs-v-pot-by-site-mature-for-obl)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_obs_v_pot_by_site_mat_for_obl",
                                      aoh_run_date,".pdf"))
```

### Range size

(ref:caption-aoh-by-range) AOH by range size.

```{r aoh-by-range, fig.cap = '(ref:caption-aoh-by-range)'}
knitr::include_graphics(path = paste0(p_plots, "aoh/trend_summary_by_range_side",
                                      aoh_run_date,".pdf"))
```


## Bonus Figures (TBD)

Reproducing all plots, with passage areas included.

1. (TBD) A summary plot showing Winners/Losers bar chart across different categories, mature forest obligates, range size.
2. (TBD) Results by season (resident, breeding, wintering) How many species have diverging trends across habitats? Perhaps some may be winners in breeding habitat, but losers for wintering habitat.
3. (TBD) Trend of overall habitat vs. AOH in abandonment periods only () vs. AOH in non-abandoned land.
  a. Might be possible to do a linear regression on this middle line. Should this be based on only abandonment periods? This is actually more complicated it needs to be the complete inverse of crop_to_abn maps. In other words, all pixels that were never abandoned plus the noncrop periods that preceded cultivation. This should be compared with the AOH in abandoned pixels only, starting with first cultivation.
4. (TBD) Winners and losers by functional group.
5. (TBD) Mapping the habitat for one species in particular, at one site, with the habitat (maybe future)

