---
title: "Implications of cropland abandonment and recultivation for biodiversity"
author:
  - Christopher L. Crawford$^1$^[Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ]
  - He Yin$^2$
  - Volker C. Radeloff$^3$
  - Yiwen Zeng$^{1}$
  - Alex Wiebe$^{4}$
  - David S. Wilcove$^{1,4}$
date: 
  $^1$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ\newline
  $^2$Department of Geography, Kent State University, Kent, OH\newline
  $^3$SILVIS Lab, Department of Forest & Wildlife Ecology, University of Wisconsin - Madison, Madison, WI\newline
  $^4$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ\newline
  \newline
  \today
output:
  bookdown::word_document2:
    reference_docx: /Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx
    number_sections: false
  bookdown::pdf_document2:
    toc: false
    keep_tex: true
    includes:
      in_header: "preamble.tex"
editor_options:
  chunk_output_type: console
bibliography: [/Users/christophercrawford/work/library/library.bib, packages.bib]
link-citations: yes
csl: science-advances.csl
urlcolor: blue
linkcolor: black
citecolor: black
---

```{r load-packages-functions, eval = TRUE, echo = FALSE, include = FALSE}
source("/Users/christophercrawford/work/projects/biodiversity_abn/scripts/0_start.R")
```

```{r preview-chapter, eval=FALSE, include = FALSE}
rmarkdown::render("scripts/Crawford_abandonment_MS+SI.Rmd",
                  output_file = 
                    paste0(p_proj, "writing/previews/", "biodiv_abn_draft",
                           format(Sys.Date(), ("_%Y_%m_%d")), ".docx"))

# Save final PDF with SI for submission:
rmarkdown::render("scripts/Crawford_biodiv_abn_MS+SI.Rmd",
                  output_file = paste0(p_proj, "writing/", "Crawford_biodiv_abn_MS+SI.pdf"),
                  output_format = "bookdown::pdf_document2")

# # Save html for troubleshooting:
# rmarkdown::render("scripts/abn_trajectories_ms.Rmd",
#                   output_file = paste0(p_proj, "writing/previews/", "abn_duration_draft_w_SI.html"),
#                   output_format = "bookdown::html_document2")

# bookdown::preview_chapter(
#   input = "abn_trajectories_ms.Rmd",
#   output_format = "bookdown::word_document2",
#   output_file = paste0("abn_duration_draft", format(Sys.Date(), ("_%Y_%m_%d")), ".docx"),
#   output_dir = paste0(p_proj, "writing/previews"),
#   output_options = 
#     list(reference_docx = "/Users/christophercrawford/work/projects/writing/scripts/word_style_ref.docx"),
#   preview = TRUE)
```

```{r set-chunk-opts, include = FALSE}
# set default chunk options
knitr::opts_chunk$set(
  out.width = "1\\textwidth",  # if compiling a pdf, include this option
  # comment = NA,
  # fig.width=6, fig.height=6, 
  echo = FALSE # the code itself will NOT be printed with results
  )

options(knitr.table.format = function() {
  if (knitr::is_latex_output())
    "latex" else "simple"
})
```

```{r produce-package-bib, include = FALSE, echo = FALSE}
# needed_packages
# github_packages
knitr::write_bib(x = c(needed_packages, github_packages), file = 'packages.bib')
# knitr::write_bib(x = c("car"), file = 'extra.bib')

# citation(package = needed_packages[1])

# knitr::write_bib(x = c(needed_packages[c(14)]), file = 'packages_top.bib')
# knitr::write_bib(x = c(needed_packages[3]), file = 'terra.bib')

# 
# needed_packages[1]
# packageDescription(needed_packages[1])
# 
# tinytex::tlmgr_install(
#   # pkgs = "ragged2e"
#   # pkgs = "standalone"
#   pkgs = "preview"
# )
```

```{r load-files, include = FALSE}

# set run:
run_label <- "_2021_03_13"

# Additional paths --------------------------------------------------- #
p_plots <- "/Users/christophercrawford/work/projects/abandonment_trajectories/output/plots/"
p_derived2 <- paste0(p_dat_derived, run_label, "/derived_data/")

# load general spatial data on site locations --------------------------------------------------- #
site_df <- read.csv(file = paste0(p_dat_derived, "site_df.csv"))
site_df_w_size <- read.csv(file = paste0(p_dat_derived, "site_df_w_size.csv"))
site_sf <- st_read(paste0(p_dat_derived, "site_sf.shp"))


# load all length data (distilled)
length_distill_df <- read_csv(file = paste0(p_derived2, 
                                            "length_distill_df", run_label, ".csv"))

# load mean length data
mean_length_df <- read_csv(file = paste0(p_derived2, "mean_length_df", run_label, ".csv"))


# load summary statistic derived from lengths
summary_stats_all_sites <- read_csv(file = paste0(p_derived2, "summary_stats_all_sites", run_label, ".csv"))
summary_stats_all_sites_pooled <- read_csv(file = paste0(p_derived2, "summary_stats_all_sites_pooled", run_label, ".csv"))

# load area summary stats:
area_summary_df <- read_csv(file = paste0(p_derived2, "area_summary_df", run_label, ".csv"))
area_recult_threshold <- read_csv(file = paste0(p_derived2, "area_recult_threshold", run_label, ".csv"))


# load stats on proportion of abandoned cropland by land cover
abn_lc_area_2017 <- read_csv(file = paste0(p_derived2, "abn_lc_area_2017", run_label, ".csv"))
abn_prop_lc_2017 <- read_csv(file = paste0(p_derived2, "abn_prop_lc_2017", run_label, ".csv"))


# load combined persistence, area, and turnover datasets:
# also known as dat_l in "4_decay_models.Rmd"

persistence_dat <- 
  # dat_l <- 
  read_csv(file = paste0(p_derived2, "persistence_dat", run_label, ".csv")) %>%
  mutate(site = as.factor(site),
         bins = as.factor(bins),
         year_abn_bins = as.factor(year_abn_bins),
         cohort = as.factor(cohort))

area_dat <- read_csv(file = paste0(p_derived2, "area_dat", run_label, ".csv"))
turnover_dat <- read_csv(file = paste0(p_derived2, "turnover_dat", run_label, ".csv"))



# load decay rate data
load(file = paste0(p_derived2, "/decay_mod_coef_dfs.rds"), 
     verbose = TRUE)

# load extrapolation data
extrapolate_df_l3 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3", run_label, ".csv"))
extrapolate_df_l3_a2 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3_a2", run_label, ".csv"))


# load data for shaanxi maps --------------------------------------------------- #

# load data for decay models --------------------------------------------------- #
fitted_df_l3 <- read_csv(file = paste0(p_derived2, "/decay_mod_fitted_df_l3", run_label, ".csv"))

load(file = paste0(p_derived2, "/decay_mod_coef_dfs.rds"), verbose = TRUE)
# contains:
# coef_df_l3, coef_df_wide_l3, mean_coef_df_l3, mean_coef_df_wide_l3, mean_coef_proj_l3, time_to_l3


# load extrapolation data
extrapolate_df_l3 <- read_csv(file = paste0(p_derived2, "/extrapolate_df_l3", run_label, ".csv"))

future_df_l3 <- read_csv(file = paste0(p_derived2, "/future_df_l3", run_label, ".csv"))

```

Preparing manuscript for *Conservation Letters*:  
- 3000 word limit
- No more than 6 figures and tables
- 40 References

# Abstract {.unnumbered}

Background/Question/Methods:

Although agricultural expansion and deforestation continues to be a dominant trend in many regions, tens of millions of hectares of croplands and pastures have simultaneously been abandoned as a result of shifting demographics, sociopolitical upheaval, environmental changes, and globalization. 
These abandoned agricultural lands increasingly represent opportunities for the restoration of ecosystems, with large potential environmental benefits in terms of carbon sequestration and increases in habitat availability for wildlife.
However, the extent to which abandonment presents an opportunity, or a threat, to biodiversity remains a topic of hot debate, due to varying species habitat requirements, threat status, and other factors.
Here, we combine a recent high-resolution annual land cover time series to estimate changes in area of habitat for bird and mammal species at 11 sites distributed across four continents.
We use annual land cover maps to identify cropland abandonment, and specifically track the age of abandoned fields.
We then calculate area of habitat (AOH) for bird and mammal species, by reconciling land cover classes to match IUCN habitat types, then excluding all land cover and elevations that are deemed unsuitable for a given species, according to the IUCN.

We 
We then explore how habitat area changed as a result of cropland abandonment by delving into differences between groups of species, looking at IUCN threat status, habitat preferences, habitat age requirements (early successional vs. mature habitat only)

We find that a significant portion of abandoned croplands are recultivated by the end of our time series, substantially curtailing the potential gains in AOH for different species.
We also find that 

Our land cover maps indicate that 
We show

We also investigate how recent abandonment and subsequent regeneration has affected landscape fragmentation. 

This time series shows that while 

agricultural abandonment , increasing levels of abandonment have raised hopes for recovery in some places.

Agricultural expansion continues to be the do

## Keywords {.unnumbered}

# Introduction {.unnumbered}

# Methods {.unnumbered}

## Abandonment maps

## Species range maps and filtering steps

## Filtering species by habitat age requirements

Species were given a binary classification indicating whether it required mature habitat (in our case, any habitat requiring at least 30 years of secondary succession to become suitable for that species), which we used to further classify our results.
Classifications were manually determined based on species accounts from the handbook of the birds of the world, and handbook of the mammals of the world.
Neotropical bird species were independently verified by expert observer Alex Wiebe.


## 

# Results {.unnumbered}

# Discussion {.unnumbered}

# Acknowledgements {.unnumbered}

## Conflict of Interest {.unnumbered}

The authors declare no conflicts of interest.

## Ethics and Funding {.unnumbered}

## Data Availability Statement {.unnumbered}

All datasets generated and analyzed for this article are archived at Zenodo (URL TBD). 
The annual land cover maps that underlie our analysis are also archived at Zenodo (URL TBD, doi:), along with derived maps of abandonment age and derived datasets summarizing area, persistence, and turnover of abandonment.
Code to replicate these analyses is available on GitHub at <https://github.com/chriscra/biodiversity_abandonment>, and is archived at Zenodo (URL TBD).

# References {.unnumbered}

BirdLife International and Handbook of the Birds of the World (2021) Bird species distribution maps of the world. Version 2021.1. Available at http://datazone.birdlife.org/species/requestdis.

Request blurb:
This IUCN Assessment data will be used for a dissertation research project designed to investigate the impact of agricultural abandonment on biodiversity. The final product will be academic journal articles, which will be shared with the IUCN science team upon completion. Please don't hesitate to get in touch if you have any questions (ccrawford@princeton.edu). Thank you! - Christopher Crawford, PhD Candidate, Princeton University.

::: {#refs}
:::

\newpage

\beginsupplement

# Supplementary information {.unnumbered}

\tableofcontents
\listoftables
\listoffigures

\newpage

# Extended Results {#results-si}

# Extended Methods

# Supplemental Figures

# Supplemental Tables
